---
title: "Price vs X"
output: html_notebook
---

Lesson 5
========================================================

### Multivariate Data
Notes:

***

### Moira Perceived Audience Size Colored by Age
Notes:

***

### Third Qualitative Variable
Notes:
```{r}
library(ggplot2)
pf <- read.csv("pseudo_facebook.tsv", sep = "\t")

```

```{r}
?stat_summary
```

```{r Third Qualitative Variable}
# Write code to create a new data frame,
# called 'pf.fc_by_age_gender', that contains
# information on each age AND gender group.

# The data frame should contain the following variables:

#    mean_friend_count,
#    median_friend_count,
#    n (the number of users in each age and gender grouping)

# Here is an example of the structure of your data frame. Your
# data values will be different. Note that if you are grouping by
# more than one variable, you will probably need to call the
# ungroup() function. 

#   age gender mean_friend_count median_friend_count    n
# 1  13 female          247.2953                 150  207
# 2  13   male          184.2342                  61  265
# 3  14 female          329.1938                 245  834
# 4  14   male          157.1204                  88 1201

ggplot(subset(pf, !is.na(gender)), aes(gender, age)) +
  geom_boxplot() + 
  stat_summary(fun.y = mean, geom = 'point', shape = 4)

ggplot(subset(pf, !is.na(gender)), aes(age, friend_count)) + 
  geom_line(aes(color = gender), stat = 'summary', fun.y = median)
```
```{r}
# Create pf.fc_by_age_gender

library(dplyr)
# Chain functions together %>%
pf.fc_by_age_gender <- pf %>%
  filter(!is.na(gender))%>%
  group_by(age, gender) %>%
  summarise(mean_friend_count = mean(friend_count),
            median_friend_count = median(friend_count),
            n = n()) %>%
  ungroup()%>%
  arrange(age)

head(pf.fc_by_age_gender)
```

***

### Plotting Conditional Summaries
Notes:

```{r Plotting Conditional Summaries}
# Create a line graph showing the
# median friend count over the ages
# for each gender. Be sure to use
# the data frame you just created,
# pf.fc_by_age_gender.
ggplot(subset(pf.fc_by_age_gender, !is.na(gender)), aes(age, median_friend_count)) +
  geom_line(aes(color = gender), stat = 'summary', fun.y = median)
```

***

### Thinking in Ratios
Notes:

***

### Wide and Long Format
Notes:

***

### Reshaping Data
Notes:

```{r}
install.packages('reshape2')
library(reshape2)
pf.fc_by_age_gender.wide <- dcast(pf.fc_by_age_gender,
                                  age ~ gender,
                                  value.var = 'median_friend_count')

head(pf.fc_by_age_gender.wide)
```


***

### Ratio Plot
Notes:

```{r Ratio Plot}
# Plot the ratio of the female to male median
# friend counts using the data frame
# pf.fc_by_age_gender.wide.

# Think about what geom you should use.
# Add a horizontal line to the plot with
# a y intercept of 1, which will be the
# base line. Look up the documentation
# for geom_hline to do that. Use the parameter
# linetype in geom_hline to make the
# line dashed.

# The linetype parameter can take the values 0-6:
# 0 = blank, 1 = solid, 2 = dashed
# 3 = dotted, 4 = dotdash, 5 = longdash
# 6 = twodash

ggplot(pf.fc_by_age_gender.wide, aes(age, female/male)) + 
  geom_line() + 
  geom_hline(yintercept = 1, alpha = 0.5, linetype = 2)
```

***

### Third Quantitative Variable
Notes:

```{r Third Quantitative Variable}
# Create a variable called year_joined
# in the pf data frame using the variable
# tenure and 2014 as the reference year.

# The variable year joined should contain the year
# that a user joined facebook.

# See the Instructor Notes for three hints if you get
# stuck. Scroll down slowly to see one hint at a time
# if you would like some guidance.

# This programming exercise WILL BE automatically graded.
pf$year_joined <- floor(2014 - pf$tenure/365)
```

***

### Cut a Variable
Notes:

```{r Cut a Variable}
# Create a new variable in the data frame
# called year_joined.bucket by using
# the cut function on the variable year_joined.

# You need to create the following buckets for the
# new variable, year_joined.bucket

#        (2004, 2009]
#        (2009, 2011]
#        (2011, 2012]
#        (2012, 2014]

# Note that a parenthesis means exclude the year and a
# bracket means include the year.

# Look up the documentation for cut or try the link
# in the Instructor Notes to accomplish this task.
summary(pf$year_joined)
table(pf$year_joined)

pf$year_joined.bucket <- cut(pf$year_joined,
                             c(2004, 2009, 2011, 2012, 2014))
table(pf$year_joined.bucket)
```

***

### Plotting it All Together
Notes:

```{r Plotting it All Together}
# Create a line graph of friend_count vs. age
# so that each year_joined.bucket is a line
# tracking the median user friend_count across
# age. This means you should have four different
# lines on your plot.

# You should subset the data to exclude the users
# whose year_joined.bucket is NA.

# If you need a hint, see the Instructor Notes.

# This assignment is not graded and
# will be marked as correct when you submit.
ggplot(subset(pf, !is.na(year_joined.bucket)), aes(age, friend_count))+
  geom_line(aes(color = year_joined.bucket), stat = 'summary', fun.y = median)
```

***

### Plot the Grand Mean
Notes:

```{r Plot the Grand Mean}
# Write code to do the following:

# (1) Add another geom_line to code below
# to plot the grand mean of the friend count vs age.

# (2) Exclude any users whose year_joined.bucket is NA.

# (3) Use a different line type for the grand mean.

# As a reminder, the parameter linetype can take the values 0-6:

# 0 = blank, 1 = solid, 2 = dashed
# 3 = dotted, 4 = dotdash, 5 = longdash
# 6 = twodash
ggplot(subset(pf, !is.na(year_joined.bucket)), aes(age, friend_count))+
  geom_line(aes(color = year_joined.bucket), 
            stat = 'summary',
            fun.y = mean) +
  geom_line(stat = 'summary', fun.y = mean, linetype = 2)
```

***

### Friending Rate
Notes:

```{r Friending Rate}
with(subset(pf, tenure >=1), summary(friend_count/tenure))
```

***

### Friendships Initiated
Notes:

What is the median friend rate? 0.2205

What is the maximum friend rate? 417.0000

```{r Friendships Initiated}
# Create a line graph of mean of friendships_initiated per day (of tenure)
# vs. tenure colored by year_joined.bucket.

# You need to make use of the variables tenure,
# friendships_initiated, and year_joined.bucket.

# You also need to subset the data to only consider user with at least
# one day of tenure.

ggplot(subset(pf, tenure >=1), aes(tenure, friendships_initiated/tenure)) +
  geom_line(aes(color = year_joined.bucket), stat = 'summary', fun.y = mean)
```

***

### Bias-Variance Tradeoff Revisited
Notes:

```{r Bias-Variance Tradeoff Revisited}

ggplot(aes(x = tenure, y = friendships_initiated / tenure),
       data = subset(pf, tenure >= 1)) +
  geom_line(aes(color = year_joined.bucket),
            stat = 'summary',
            fun.y = mean)

ggplot(aes(x = 7 * round(tenure / 7), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

ggplot(aes(x = 30 * round(tenure / 30), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

ggplot(aes(x = 90 * round(tenure / 90), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

```
```{r}
?geom_smooth
```

```{r}
ggplot(aes(x = tenure, y = friendships_initiated / tenure),
       data = subset(pf, tenure >= 1)) +
  geom_smooth(aes(color = year_joined.bucket))
```

***

### Sean's NFL Fan Sentiment Study
Notes:

***

### Introducing the Yogurt Data Set
Notes:

***

### Histograms Revisited
Notes:

```{r Histograms Revisited}
yo <- read.csv('yogurt.csv')
str(yo)

# Change the id from an int to a factor
yo$id <- factor(yo$id)
str(yo)

# Create an histogram
ggplot(yo, aes(price)) +
  geom_histogram() + 
  scale_x_continuous(breaks = seq(0,70,5))

# Let's change the binwidth to 10
ggplot(yo, aes(price)) +
  geom_histogram(binwidth = 5) + 
  scale_x_continuous(breaks = seq(0,70,5))

ggplot(yo, aes(price)) +
  geom_histogram(binwidth = 10) + 
  scale_x_continuous(breaks = seq(0,70,10))
```

***

### Number of Purchases
Notes:
```{r}
?transform
```

```{r Number of Purchases}
summary(yo)
length(unique(yo$price))
table(yo$price)

# Create a new variable called all.purchases,
# which gives the total counts of yogurt for
# each observation or household.

# One way to do this is using the transform
# function. You can look up the function transform
# and run the examples of code at the bottom of the
# documentation to figure out what it does.

# The transform function produces a data frame
# so if you use it then save the result to 'yo'!

# OR you can figure out another way to create the
# variable.

yo <- transform(yo, all.purchases = strawberry + blueberry + pina.colada + plain + mixed.berry)
summary(yo)

# Alternative 1
# yo$all.purchases <- yo$strawberry + yo$blueberry + yo$pina.colada + yo$plain + yo$mixed.berry

# Alternative 2
# with(yo, all.purchases <- strawberry + blueberry + pina.colada + plain + mixed.berry)

```

***

### Prices over Time
Notes:

```{r Prices over Time}
ggplot(yo, aes(all.purchases)) +
  geom_histogram(binwidth = 1)

ggplot(yo, aes(time, price)) +
  geom_jitter(alpha = 0.25, shape = 21, fill = I('#F79420'))
```

***

### Sampling Observations
Notes:

***

### Looking at Samples of Households

```{r Looking at Sample of Households}
set.seed(4230)
sample.ids <- sample(levels(yo$id), 16)

ggplot(subset(yo, id %in% sample.ids), aes(time, price)) + 
  facet_wrap(~id) +
  geom_line() +
  geom_point(aes(size = all.purchases), pch = 1)


set.seed(1111)
sample.ids <- sample(levels(yo$id), 16)

ggplot(subset(yo, id %in% sample.ids), aes(time, price)) + 
  facet_wrap(~id) +
  geom_line() +
  geom_point(aes(size = all.purchases), pch = 1)
```

***

### The Limits of Cross Sectional Data
Notes:

***

### Many Variables
Notes:

***

### Scatterplot Matrices
Notes:

```{r}
install.packages('GGally')
```
```{r}
library(GGally)
theme_set(theme_minimal(20))

#set the seed for reproducible results
set.seed(1836)
pf_subset <- pf[, c(2:15)]
names(pf_subset)
ggpairs(pf_subset[sample.int(nrow(pf_subset), 1000), ])
```

***

### Even More Variables
Notes:

***

### Heat Maps
Notes:

```{r}
nci <- read.table("nci.tsv")
colnames(nci) <- c(1:64)
```

```{r}
nci.long.samp <- melt(as.matrix(nci[1:200,]))
names(nci.long.samp) <- c("gene", "case", "value")
head(nci.long.samp)

ggplot(aes(y = gene, x = case, fill = value),
  data = nci.long.samp) +
  geom_tile() +
  scale_fill_gradientn(colours = colorRampPalette(c("darkblue", "red"))(100))
```


***

### Analyzing Three of More Variables
Reflection:
We have learnt how to express more than 2 variables in a plot. First of all, we segmented by gender, applying a color pattern inside the aes() function in a geom_line plot. After that, we customized some cuts in years in order to get a better distribution of friend count by age based in the tenure of people.

We have also learnt that some datasets have limitations as they don't allow you to inspect the behavior of a particular sample over the time, as we did with the Yogurt's dataset, where we could inspect a user (with an unique id) and its consumer behavior over the time and the influence of yogurt prices on it.

Finally, we found a first approach to get the big picture of the relation between multiple variable with the scatterplot matrix and the heat map. This could make it easier to pick at first sight the relevant cases you may want to inspect in-depth

***


## Problem set: Explore Many Variables
```{r}
library(ggplot2)
data(diamonds)
summary(diamonds)
```
```{r}
quantile(diamonds$price, .90)
```


### Problem 1: Price Histograms with Facet and Color
```{r}
# Create a histogram of diamond prices.
# Facet the histogram by diamond color
# and use cut to color the histogram bars.

# The plot should look something like this.
# http://i.imgur.com/b5xyrOu.jpg

# Note: In the link, a color palette of type
# 'qual' was used to color the histogram using
# scale_fill_brewer(type = 'qual')

# This assignment is not graded and
# will be marked as correct when you submit.

ggplot(diamonds, aes(price)) +
  geom_histogram(aes(fill = cut)) +
  facet_wrap(~color, ncol = 2) +
  scale_x_continuous(trans = 'log10') +
  scale_fill_brewer(type = 'qual') + 
  labs(title = "Diamond Prices by Color&Cut") 
```
## Problem 2: Price vs. Table Colored by Cut
```{r}
# Create a scatterplot of diamond price vs.
# table and color the points by the cut of
# the diamond.

# The plot should look something like this.
# http://i.imgur.com/rQF9jQr.jpg

# Note: In the link, a color palette of type
# 'qual' was used to color the scatterplot using
# scale_color_brewer(type = 'qual')

# We calculate the range and will get rid of some outliers (6 items in this case)
range(diamonds$table)

ggplot(diamonds, aes(table, price)) +
  geom_jitter(aes(color = cut), alpha = 0.2) +
  scale_color_brewer(type = 'qual') +
  scale_x_continuous(limits = c(50,80), breaks = seq(43, 95, 2))
```

## Problem 3: Typical Table Value
Question:
What is the typical range for the majority of diamonds of ideal cut?

53-57

What is the typical range for the majority of diamonds of premium cut?

58-52

(Use the graph that you created from the previous exercise to see the answer. You do not need to run summaries.)

## Problem 4: Price vs. Volume and Diamond Clarity
```{r}
# Create a scatterplot of diamond price vs.
# volume (x * y * z) and color the points by
# the clarity of diamonds. Use scale on the y-axis
# to take the log10 of price. You should also
# omit the top 1% of diamond volumes from the plot.

# Note: Volume is a very rough approximation of
# a diamond's actual volume.

# The plot should look something like this.
# http://i.imgur.com/excUpea.jpg

# Note: In the link, a color palette of type
# 'div' was used to color the scatterplot using
# scale_color_brewer(type = 'div')
diamonds <- transform(diamonds, volume = x * y * z)
diamonds$volume[diamonds$volume == 0] <- NA

ggplot(diamonds, aes(volume, price)) + 
  geom_jitter(aes(color = clarity), alpha = 0.4) +
  scale_x_continuous(limits = c(0, quantile(diamonds$volume, 0.99, na.rm=TRUE))) +
  scale_y_continuous(trans = "log10") +
  scale_color_brewer(type = 'div')
```


## Problem 5: Proportion of Friendships Initiated
```{r}
# Many interesting variables are derived from two or more others.
# For example, we might wonder how much of a person's network on
# a service like Facebook the user actively initiated. Two users
# with the same degree (or number of friends) might be very
# different if one initiated most of those connections on the
# service, while the other initiated very few. So it could be
# useful to consider this proportion of existing friendships that
# the user initiated. This might be a good predictor of how active
# a user is compared with their peers, or other traits, such as
# personality (i.e., is this person an extrovert?).

# Your task is to create a new variable called 'prop_initiated'
# in the Pseudo-Facebook data set. The variable should contain
# the proportion of friendships that the user initiated.

# This programming assignment WILL BE automatically graded.

pf <- read.csv('pseudo_facebook.tsv', sep = "\t")

pf <- transform(pf, prop_initiated = friendships_initiated/friend_count)
```


## Problem 6: prop_initiated vs. tenure
```{r}
?cut
```

```{r}
# Create a line graph of the median proportion of
# friendships initiated ('prop_initiated') vs.
# tenure and color the line segment by
# year_joined.bucket.

# Recall, we created year_joined.bucket in Lesson 5
# by first creating year_joined from the variable tenure.
# Then, we used the cut function on year_joined to create
# four bins or cohorts of users.

# (2004, 2009]
# (2009, 2011]
# (2011, 2012]
# (2012, 2014]

# The plot should look something like this.
# http://i.imgur.com/vNjPtDh.jpg
# OR this
# http://i.imgur.com/IBN1ufQ.jpg


library(dplyr)
# Add year_joined column
pf$year_joined <- floor(2014 - pf$tenure/365)
# Alternative: pf$year_joined <- 2014-ceiling(pf$tenure/365)

# Add bucket column
pf$year_joined.bucket <- cut(pf$year_joined, c(2004, 2009, 2011, 2012, 2014), right = TRUE) 

ggplot(pf.prop_by_tenure, aes(tenure, median_prop_initiated)) +
  geom_line(aes(color = year_joined.bucket), stat = 'summary', fun.y = median, na.rm = TRUE)
```

## Problem 7: Smoothing prop_initiated vs. tenure
```{r}
# Smooth the last plot you created of
# of prop_initiated vs tenure colored by
# year_joined.bucket. You can bin together ranges
# of tenure or add a smoother to the plot.

# There won't be a solution image for this exercise.
# You will answer some questions about your plot in
# the next two exercises.
ggplot(pf, aes(tenure, prop_initiated)) +
  geom_smooth(aes(color = year_joined.bucket), na.rm = TRUE)

ggplot(pf, aes(x=25*round(tenure/25), y=prop_initiated, color=year_joined.bucket)) +
  geom_line(stat='summary', fun.y=median, na.rm=TRUE)

```
## Problem 8: Greatest prop_initiated group
On average, which group initiated the greatest proportion of its Facebook friendships? The plot with the smoother that you created in the last exercise can help you answer this question

* People who joined after 2012

## Problem 9: Largest Group Mean prop_initiated

```{r}
mean(pf$prop_initiated[pf$year_joined.bucket=='(2012,2014]'], na.rm=TRUE)
```

## Problem 10: Price/Carat Binned, Faceted, & Colored
```{r}
# Create a scatter plot of the price/carat ratio
# of diamonds. The variable x should be
# assigned to cut. The points should be colored
# by diamond color, and the plot should be
# faceted by clarity.

# The plot should look something like this.
# http://i.imgur.com/YzbWkHT.jpg.

# Note: In the link, a color palette of type
# 'div' was used to color the histogram using
# scale_color_brewer(type = 'div')

ggplot(diamonds, aes(cut, price/carat)) +
  geom_jitter(aes(color = color), alpha = 0.4) + 
  facet_wrap(~clarity) +
  scale_color_brewer(type = 'div')

```

Problem 11: Gapminder Multivariate Analysis
```{r}
# The Gapminder website contains over 500 data sets with information about
# the world's population. Your task is to continue the investigation you did at the
# end of Problem Set 4 or you can start fresh and choose a different
# data set from Gapminder.

# If you're feeling adventurous or want to try some data munging see if you can
# find a data set or scrape one from the web.

# In your investigation, examine 3 or more variables and create 2-5 plots that make
# use of the techniques from Lesson 5.

# You can find a link to the Gapminder website in the Instructor Notes.

# Once you've completed your investigation, create a post in the discussions that includes:
#       1. the variable(s) you investigated, your observations, and any summary statistics
#       2. snippets of code that created the plots
#       3. links to the images of your plots

```


Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

