---
title: "Red & White Wine Analysis"
output: html_notebook
---
TITLE by YOUR_NAME_HERE
========================================================

> **Tip**: You will see quoted sections like this throughout the template to
help you construct your report. Make sure that you remove these notes before
you finish and submit your project!

> **Tip**: One of the requirements of this project is that your code follows
good formatting techniques, including limiting your lines to 80 characters or
less. If you're using RStudio, go into Preferences \> Code \> Display to set up
a margin line to help you keep track of this guideline!

###1. Citation

This dataset is public available for research. The details are described in [Cortez et al., 2009].

P. Cortez, A. Cerdeira, F. Almeida, T. Matos and J. Reis. Modeling wine preferences by data mining from physicochemical properties. In Decision Support Systems, Elsevier, 47(4):547-553. ISSN: 0167-9236.

Available at: [@Elsevier] http://dx.doi.org/10.1016/j.dss.2009.05.016 [Pre-press (pdf)] http://www3.dsi.uminho.pt/pcortez/winequality09.pdf [bib] http://www3.dsi.uminho.pt/pcortez/dss09.bib

###2. About dataset

In the above reference, two datasets were created, using red and white wine samples. The inputs include objective tests (e.g. PH values) and the output is based on sensory data (median of at least 3 evaluations made by wine experts). Each expert graded the wine quality between 0 (very bad) and 10 (very excellent).

The two datasets are related to red and white variants of the Portuguese "Vinho Verde" wine. For more details, consult: http://www.vinhoverde.pt/en/ or the reference [Cortez et al., 2009]. Due to privacy and logistic issues, only physicochemical (inputs) and sensory (the output) variables are available (e.g. there is no data about grape types, wine brand, wine selling price, etc.).

###3. Number of Instances: red wine - 1599; white wine - 4898.

###4. Number of Attributes: 11 + output attribute

###5. Attribute information:

   For more information, read [Cortez et al., 2009].

   Input variables (based on physicochemical tests):
   1 - fixed acidity (tartaric acid - g / dm^3)
   2 - volatile acidity (acetic acid - g / dm^3)
   3 - citric acid (g / dm^3)
   4 - residual sugar (g / dm^3)
   5 - chlorides (sodium chloride - g / dm^3
   6 - free sulfur dioxide (mg / dm^3)
   7 - total sulfur dioxide (mg / dm^3)
   8 - density (g / cm^3)
   9 - pH
   10 - sulphates (potassium sulphate - g / dm3)
   11 - alcohol (% by volume)
   Output variable (based on sensory data): 
   12 - quality (score between 0 and 10)
###6. Description of attributes:

1 - fixed acidity: most acids involved with wine or fixed or nonvolatile (do not evaporate readily)

2 - volatile acidity: the amount of acetic acid in wine, which at too high of levels can lead to an unpleasant, vinegar taste

3 - citric acid: found in small quantities, citric acid can add 'freshness' and flavor to wines

4 - residual sugar: the amount of sugar remaining after fermentation stops, it's rare to find wines with less than 1 gram/liter and wines with greater than 45 grams/liter are considered sweet

5 - chlorides: the amount of salt in the wine

6 - free sulfur dioxide: the free form of SO2 exists in equilibrium between molecular SO2 (as a dissolved gas) and bisulfite ion; it prevents microbial growth and the oxidation of wine

7 - total sulfur dioxide: amount of free and bound forms of S02; in low concentrations, SO2 is mostly undetectable in wine, but at free SO2 concentrations over 50 ppm, SO2 becomes evident in the nose and taste of wine

8 - density: the density of water is close to that of water depending on the percent alcohol and sugar content

9 - pH: describes how acidic or basic a wine is on a scale from 0 (very acidic) to 14 (very basic); most wines are between 3-4 on the pH scale

10 - sulphates: a wine additive which can contribute to sulfur dioxide gas (S02) levels, wich acts as an antimicrobial and antioxidant

11 - alcohol: the percent alcohol content of the wine

Output variable (based on sensory data): 12 - quality (score between 0 and 10)



```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Load all of the packages that you end up using in your analysis in this code
# chunk.

# Notice that the parameter "echo" was set to FALSE for this code chunk. This
# prevents the code from displaying in the knitted HTML output. You should set
# echo=FALSE for all code chunks in your file, unless it makes sense for your
# report to show the code that generated a particular plot.

# The other parameters for "message" and "warning" should also be set to FALSE
# for other code chunks once you have verified that each plot comes out as you
# want it to. This will clean up the flow of your report.

library(ggplot2)
library(dplyr)
```

```{r}
# Load the Data
getwd()
setwd("C:/Users/Serg/Documents/Aprendiendo/Data Analysis/Module 5 - Exploratory Data Analysis (EDA)/Proyecto final")

# read the .csv files and create a dataframe for each kind of wine
red_wine <- read.csv('wineQualityReds.csv')
white_wine <- read.csv('wineQualityWhites.csv')

# add a column called "typè" to categorize each type of wine
red_wine['type'] <- 'red'
white_wine['type'] <- 'white'

# bind both datasets
wine_data <- rbind(red_wine, white_wine)
```

> **Tip**: Before you create any plots, it is a good idea to provide a short
introduction into the dataset that you are planning to explore. Replace this
quoted text with that general information!

```{r}
dim(wine_data)
head(wine_data)
summary(wine_data)
```

Observations from the Summary
Mean residual sugar level is 5.4 g/l, but there is a sample of very sweet wine with 65.8 g/l (an outlier). Mean free sulfur dioxide is 30.5 ppm. Max value is 289 which is quite high as 75% is 41 ppm. PH of wine is within range from 2.7 till 4, mean 3.2. There is no basic wines in this dataset (no high pH levels). Alcohol: lightest wine is 8%, strongest is 14.9. Minimum quality mark is 3, mean 5.8, highest is 9.

```{r}
head(subset(wine_data, wine_data['type'] == 'red'))
dim(subset(wine_data, wine_data['type'] == 'red'))
summary(subset(wine_data, wine_data['type'] == 'red'))
```

```{r}
head(subset(wine_data, wine_data['type'] == 'white'))
dim(subset(wine_data, wine_data['type'] == 'white'))
summary(subset(wine_data, wine_data['type'] == 'white'))
```


# Univariate Plots Section

> **Tip**: In this section, you should perform some preliminary exploration of
your dataset. Run some summaries of the data and create univariate plots to
understand the structure of the individual variables in your dataset. Don't
forget to add a comment after each plot or closely-related group of plots!
There should be multiple code chunks and text sections; the first one below is
just to help you get started.

```{r}
ggplot(wine_data, aes(type, residual.sugar))+
  geom_boxplot()+ 
  stat_summary(fun.y=mean, color="orange", geom="point", 
               shape=18, size=3)
```


```{r}
summary(wine_data$quality)
table(wine_data$quality)
qplot(quality, data = wine_data, fill = type, binwidth = 1)+
    scale_x_continuous(breaks = seq(3,10,1), lim = c(3,10)) +
    scale_y_sqrt()

ggplot(wine_data, aes(quality))+
  geom_histogram(aes(y=..density..), breaks = seq(3,10,1))+
  facet_wrap(~type, ncol = 1 ) +
  geom_vline(data=filter(wine_data,type == "red"),aes(xintercept = mean(quality),group = type),col='blue',size=0.4, linetype = "dashed")+
    geom_vline(data=filter(wine_data, type == "white"),aes(xintercept = mean(quality),group = type),col='blue',size=0.4, linetype = "dashed")
```

  

We know that number of observations for red and white are different in out dataset, but still we can see that for both colors it's normal distribution with almost the same picks at 5 and 6 quality point.



```{r}
summary(wine_data$alcohol)
ggplot(wine_data, aes(alcohol)) + 
  geom_histogram() +
  facet_grid(type~.)
qplot(alcohol, data = wine_data, fill = type, binwidth = 0.5) +
    scale_x_continuous(breaks = seq(8,15,0.5), lim = c(8,15))
```
Alcohol level distribution looks skewed. Again, red wine sample is smaller but it gives the same pattern of alcohol level distribution as while wines. Most frequently wines have 9.5%, mean is 10.49% of alcohol.



```{r}
summary(wine_data$density)

ggplot(wine_data, aes(density)) + 
  geom_histogram() +
  facet_grid(type~.)
  
qplot(density, data = wine_data, fill = type, binwidth = 0.0002) +
    scale_x_log10(lim = c(min(wine_data$density), 1.00370), 
                  breaks = seq(min(wine_data$density), 1.00370, 0.002))
```

Looking at 'table' summary we see that there are two outliers: 1.0103 and 1.03898. To see the distribution of density clearer I used log10 and limited the data. Now we can see that density distribution of white wine is bimodal and of red wine is normal.



```{r}
summary(wine_data$volatile.acidity)
ggplot(wine_data, aes(volatile.acidity)) + 
  geom_histogram() +
  facet_grid(type~.)
qplot(volatile.acidity, data = wine_data, fill = type, binwidth = 0.001) +
    scale_x_log10(breaks = seq(min(wine_data$volatile.acidity), 
                               max(wine_data$volatile.acidity), 0.1)) +
  scale_y_log10()
```

Volatile acidity has normal distribution. I also suppose that more acetic wines have worse marks because high acidity can lead to unpleasant taste.


```{r}
summary(wine_data $chlorides)

ggplot(wine_data, aes(chlorides)) + 
  geom_histogram() +
  facet_grid(type~.)

qplot(chlorides, data = wine_data, fill = type, binwidth = 0.01) +
    scale_x_log10(breaks = seq(min(wine_data$chlorides), quartile(wine_data$chlorides), 0.9))
```

Chlorides distribution initially is skewed so I used log10 to see the distribution clearer.

```{r}
ggplot(wine_data, aes(chlorides)) + 
  geom_histogram() +
  facet_grid(type~quality)
```


> **Tip**: Make sure that you leave a blank line between the start / end of
each code block and the end / start of your Markdown text so that it is
formatted nicely in the knitted text. Note as well that text on consecutive
lines is treated as a single space. Make sure you have a blank line between
your paragraphs so that they too are formatted for easy readability.

# Univariate Analysis

> **Tip**: Now that you've completed your univariate explorations, it's time to
reflect on and summarize what you've found. Use the questions below to help you
gather your observations and add your own if you have other thoughts!

### What is the structure of your dataset?

### What is/are the main feature(s) of interest in your dataset?

### What other features in the dataset do you think will help support your \
investigation into your feature(s) of interest?

### Did you create any new variables from existing variables in the dataset?

### Of the features you investigated, were there any unusual distributions? \
Did you perform any operations on the data to tidy, adjust, or change the form \
of the data? If so, why did you do this?


# Bivariate Plots Section

> **Tip**: Based on what you saw in the univariate plots, what relationships
between variables might be interesting to look at in this section? Don't limit
yourself to relationships between a main output feature and one of the
supporting variables. Try to look at relationships between supporting variables
as well.

```{r echo=FALSE, Bivariate_Plots}
qplot(quality, data = wine_data, binwidth = 1, color = type, geom = "density") + 
    scale_x_continuous(breaks = seq(3, 9, 1))
```
In out sample we have almost the same amount of red and white wines with quality '3', '4' and '9', more red wines with quality '5' and more white wines with quality "6", "7" and "8".




```{r}
ggpairs(wine_data)

```

I checked correlation between all the variables in wine dataset.

We can see some correlation in pairs like:

alcohol vs. density
fixed.acidity vs. density
residual.sugar vs total.sulfur.dioxide
residual.sugar vs. density
residual.sugar vs. alcohol
chlorides vs. density
chlorides vs. sulphates
quality vs. alcohol


Compare objective parameters of wines

```{r}
# use function for plotting with ggplot for simplicity of code
f <- function(wine_data, x, y, z, opts=NULL) {
  ggplot(wine_data, aes_string(x = x, y = y, color = z)) +
   geom_point(alpha = 1/5, position = position_jitter(h = 0), size = 2) +
   geom_smooth(method = 'lm')
}

# density vs. alcohol plot
p <- f(wine_data, "density", "alcohol", "type")
p + coord_cartesian(xlim=c(min(wine_data$density),1.005), ylim=c(8,15))

# density vs. fixed.acidity plot
p <- f(wine_data, "density", "fixed.acidity", "type")
p + coord_cartesian(xlim=c(min(wine_data$density),1.005))

# residual.sugar vs. total.sulfur.dioxide
p <- f(wine_data, "residual.sugar", "total.sulfur.dioxide", "type")
p + scale_x_log10() +
    coord_cartesian(xlim=c(min(wine_data$residual.sugar),30), 
                    ylim=c(min(wine_data$total.sulfur.dioxide), 350))

# residual.sugar vs. density
p <- f(wine_data, "residual.sugar", "density", "type")
p + coord_cartesian(xlim=c(min(wine_data$residual.sugar),25), 
                    ylim=c(min(wine_data$density), 1.005))

# residual.sugar vs. alcohol
p <- f(wine_data, "residual.sugar", "alcohol", "type")
p + coord_cartesian(xlim=c(min(wine_data$residual.sugar),25), 
                    ylim=c(min(wine_data$alcohol), 15))

# chlorides vs. density
p <- f(wine_data, "chlorides", "density", "type")
p + scale_x_log10() +
    coord_cartesian(ylim=c(min(wine_data$density), 1.005))

# chlorides vs. sulphates
p <- f(wine_data, "chlorides", "sulphates", "type")
p + scale_x_log10() +
    coord_cartesian(ylim=c(min(wine_data$sulphates), 1))
```

After checking correlated pairs I noticed that red and white wine behave different in some graphs. Means for red and white wine correlation can be significantly different.



```{r}
                                              RED      |     WHITE
* alcohol vs. density                    :    strong c.  :   strong c.
* fixed.acidity vs. density              :    strong c.  :     no c.
* residual.sugar vs total.sulfur.dioxide :     weak      :    weak c.
* residual.sugar vs. density             :    strong c.  :   strong c. 
* residual.sugar vs. alcohol             :     no c.     :   strong c.
* chlorides vs. density                  :    strong c.  :   strong c.
* chlorides vs. sulphates                :    strong c.  :     no c.
```

Compare Quality vs. Alcohol

```{r}
ggplot(data = wine_data, aes(y = alcohol, x = quality)) +
   geom_point(alpha = 1/4, position = position_jitter(h = 0), size = 4) +
   geom_smooth(method = 'lm') +
   facet_wrap(~ type)
```

My idea that volatile acidity affects the quality of wine is not correct. The only objective wine parameter which has correlation with quality is alcohol.

Significantly Different Parameters for Red and White Wine

Fixed Acidity vs. Color

```{r}
qplot(x = type, y = fixed.acidity, data = wine_data, geom = "boxplot")

```

Volatile Acidity vs. Color

```{r}
qplot(x = type, y = volatile.acidity, data = wine_data, geom = "boxplot")

```

Residual Sugar vs. Color

```{r}
qplot(x = type, y = residual.sugar, data = wine_data, geom = "boxplot")

```

Total Sulfur Dioxide vs. Color

```{r}
qplot(x = type, y = total.sulfur.dioxide, data = wine_data, geom = "boxplot")

```

These are parameters which depend a lot on the color of wine.



Quality of Red and White Wines

```{r}
qplot(x = type, y = quality, data = wine_data, geom = "boxplot")

```
Mean and 25% & 75% quintiles are similar for red and white wines.




Exploring Wine Parameters and Quality
Chlorides and Sulphates
```{r}
# converting 'quality' vector into factor variable
wine_data$quality <- as.factor(wine_data$quality)

ggplot(aes(x = chlorides, y = sulphates, color = quality), data = wine_data) +
    facet_wrap(~type) + 
    geom_point(size = 3, alpha = 1/4) +
    scale_color_identity(guide = 'legend') +
    ylim(min(wine_data$sulphates), quantile(wine_data$sulphates, 0.95)) +
    xlim(min(wine_data$chlorides), quantile(wine_data$chlorides, 0.95)) 
```

Sulfates and chlorides of white wine are spread out more than red wine. Most frequent quality levels for both colors are 4, 5, 6 and 7.



Fixed Acidity and Volatile Acidity

```{r}
ggplot(aes(x = fixed.acidity, 
           y = volatile.acidity, 
           color = quality), 
       data = wine_data) +
    facet_wrap(~type) + 
    geom_point(size = 3, alpha = 1/4) +
    scale_color_identity(guide = 'legend') +
    ylim(min(wine_data$volatile.acidity), 
         quantile(wine_data$volatile.acidity, 0.99)) +
    xlim(min(wine_data$fixed.acidity), 
         quantile(wine_data$fixed.acidity, 0.99)) 
```

Red wine of quality 5 has fixed acidity between 6 - 12, white wine = 5 - 10. White wine samples of quality '6' are highly concentrated around volatile acidity of 0.2 and fixed acidity of 6-7.



Free Sulfur Dioxide and Total Sulfur Dioxide

```{r}
ggplot(aes(x = free.sulfur.dioxide, 
           y = total.sulfur.dioxide, 
           color = quality), 
       data = wine_data) +
    facet_wrap(~type) + 
    geom_point(size = 3, alpha = 1/4) +
    scale_color_identity(guide = 'legend') +
    ylim(min(wine_data$total.sulfur.dioxide), 
         quantile(wine_data$total.sulfur.dioxide, 0.95)) +
    xlim(min(wine_data$free.sulfur.dioxide), 
         quantile(wine_data$free.sulfur.dioxide, 0.95)) 
```

We can see clear positive correlation between total sulfur dioxide and free sulfur dioxide for red and white wine. Total sulfur dioxide of white wine has higher values mostly because of wine with quality '5'.



PH and Alcohol

```{r}
ggplot(aes(x = pH, y = alcohol, color = quality), data = wine_data) +
    facet_wrap(~type) + 
    geom_point(size = 3, alpha = 1/4) +
    scale_color_identity(guide = 'legend') +
    ylim(min(wine_data$alcohol), quantile(wine_data$alcohol, 0.95)) +
    xlim(min(wine_data$pH), quantile(wine_data$pH, 0.95)) 
```

Levels of alcohol for both wine is quite the same, but minimum level of pH for white wine is 2.9, for red wine is 3.1. Only red wine with quality '5' has pH level around 3.



Citric Acid and Alcohol

```{r}
ggplot(aes(x = citric.acid, y = alcohol, color = quality), 
       data = wine_data) +
    facet_wrap(~type) + 
    geom_point(size = 3, alpha = 1/4) +
    scale_color_identity(guide = 'legend') +
    ylim(min(wine_data$alcohol), quantile(wine_data$alcohol, 0.95)) +
    xlim(min(wine_data$citric.acid), quantile(wine_data$citric.acid, 0.95)) 
```
In these plots we can notice that most of red wine is clustered around where citric acid is 0 - 0.2, for white wine this is true for citric acid level 0.2 - 0.4.






# Bivariate Analysis

> **Tip**: As before, summarize what you found in your bivariate explorations
here. Use the questions below to guide your discussion.

### Talk about some of the relationships you observed in this part of the \
investigation. How did the feature(s) of interest vary with other features in \
the dataset?

### Did you observe any interesting relationships between the other features \
(not the main feature(s) of interest)?

### What was the strongest relationship you found?


# Multivariate Plots Section

> **Tip**: Now it's time to put everything together. Based on what you found in
the bivariate plots section, create a few multivariate plots to investigate
more complex interactions between variables. Make sure that the plots that you
create here are justified by the plots you explored in the previous section. If
you plan on creating any mathematical models, this is the section where you
will do that.

```{r echo=FALSE, Multivariate_Plots}

```

# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. Were there features that strengthened each other in terms of \
looking at your feature(s) of interest?

### Were there any interesting or surprising interactions between features?

### OPTIONAL: Did you create any models with your dataset? Discuss the \
strengths and limitations of your model.

------

# Final Plots and Summary

> **Tip**: You've done a lot of exploration and have built up an understanding
of the structure of and relationships between the variables in your dataset.
Here, you will select three plots from all of your previous exploration to
present here as a summary of some of your most interesting findings. Make sure
that you have refined your selected plots for good titling, axis labels (with
units), and good aesthetic choices (e.g. color, transparency). After each plot,
make sure you justify why you chose each plot by describing what it shows.

### Plot One: Quality of Wine

```{r echo=FALSE, Plot_One}
summary(wine_data$quality)
table(wine_data$quality)
qplot(as.numeric(as.character(quality)), 
      data = wine_data, 
      fill = type, 
      binwidth = 1, 
      origin = - 0.5, 
      main = "Quality of Red and White Wine") +
    scale_x_continuous(breaks = seq(2,10,1), lim = c(2,10)) +
    scale_y_sqrt(breaks = seq(0,5600,500)) +
    xlab('Quality') +
    ylab('Quantity')
```

### Description One
Let's again look at distribution of wine by color. We may say that most frequent quality levels are 5 and 6 for both wine color.



### Plot Two: Correlation Between Objective Parameters

```{r echo=FALSE, Plot_Two}
ggplot(data = wine_data, 
       aes(x = density, y = alcohol, color = type)) +
   geom_point(alpha = 1/6, position = position_jitter(h = 0), size = 3) +
   geom_smooth(method = 'lm') +
   coord_cartesian(xlim=c(min(wine_data$density),1.005), ylim=c(8,15)) +
   xlab('Density') +
   ylab('Alcohol') +
   ggtitle('Density vs. Alcohol correlation by Color')
```

### Description Two
Density and Alcohol show the strongest correlation among all wine parameters and it equals to -0.687 for both red and white wine. Red wine in average are stronger than white wine. In this dataset wine with less alcohol percentage are mostly white and red wine mostly has more alcohol percentage.

```{r}
ggplot(data = wine_data, 
       aes(x = density, y = alcohol, color = factor(quality))) +
   geom_point(alpha = 1/2, position = position_jitter(h = 0), size = 2) +
   coord_cartesian(xlim=c(min(wine_data$density),1.005), ylim=c(8,15)) +
   scale_color_brewer(type='qual') +
   xlab('Density') +
   ylab('Alcohol') +
   ggtitle('Density vs. Alcohol correlation by Quality')
```

```{r}
ggplot(data = wine_data, 
       aes(x = density, y = alcohol) )+
   facet_wrap( ~ quality) +
   geom_boxplot() +
   xlab('Density') +
   ylab('Alcohol') +
   ggtitle('Density vs. Alcohol correlation by Quality')
```
### Description 
Wine with high alcohol percentage has quality level 7, wine with less alcohol percentage is quality level 5. Wine with quality levels 6 and 8 have various combinations of alcohol and density.


### Plot Three: Quality vs. Alcohol

```{r echo=FALSE, Plot_Three}
ggplot(data = wine_data, aes(y = alcohol, x = quality)) +
   geom_boxplot() +
   geom_smooth(method = 'lm') +
   facet_wrap(~ type) +
   xlab('Quality') +
   ylab('Alcohol') +
   ggtitle('How Alcohol Level Affects Wine Quality')
```

### Description Three
Alcohol level and Quality have correlation value of 0.4. This is the strongest correlation we have found between an objective wine parameter and wine quality. 0.4 is not a high correlation level so we cannot use alcohol as a parameter for quality prediction.


------

# Reflection

> **Tip**: Here's the final step! Reflect on the exploration you performed and
the insights you found. What were some of the struggles that you went through?
What went well? What was surprising? Make sure you include an insight into
future work that could be done with the dataset.

> **Tip**: Don't forget to remove this, and the other **Tip** sections before
saving your final work and knitting the final report!

Analyzing the data we can come up the following conclusion:

1. When alcohol percentage decreases, density grows.
2. In general alcohol level of red wine is higher than alcohol level of white wine.
3. When fixed acidity increases density of red wine increases as well. White wine almost doesn't show any correlation.
4. Total sulfur dioxide and level of residual sugar are positively correlated. Correlation shows higher value with white wine.
5. White wine density and residual sugar level have positive correlation.
6. Alcohol level of white wine decreases with the growth of residual sugar level.
7. Wine with high alcohol percentage has quality level 7, wine with less alcohol percentage is quality level 5.
8. Mostly frequent quality levels of red and white wine are 5 and 6.

According to my investigation I may conclude that experts' decisions on wine quality levels are based on their personal testes or could depend on other variables like year of production, grape types, wine brand etc. as only one variable (alcohol level) has correlation with quality of wine.

For future exploration of this data I would pick one category of wine (for example, quality level 3-4, 5-7, 8-9) to look at the patterns which can appear in each of these three buckets. I also would normalize data because we have more white wine than red wine.