---
title: "Red & White Wine Analysis"
output: html_document
---
by Sergio Marfull 
========================================================

______________________

## About datasets

### Description

This datasets are public available for research. The details are described in 
[Cortez et al., 2009].

P. Cortez, A. Cerdeira, F. Almeida, T. Matos and J. Reis. *"Modeling wine 
preferences by data mining from physicochemical properties"*. In Decision 
Support Systems, Elsevier, 47(4):547-553. ISSN: 0167-9236.

Available at: 

* [Elsevier](http://dx.doi.org/10.1016/j.dss.2009.05.016)

* [Pre-press (pdf)](http://www3.dsi.uminho.pt/pcortez/winequality09.pdf)

* [bib](http://www3.dsi.uminho.pt/pcortez/dss09.bib)


### Sample size: 
* **Red wine**: 1,599 samples; 
* **White wine**: 4,898 samples.

### Attributes:

These datasets consist of 11 attributes and one output variable (below)

* **Fixed acidity** (tartaric acid - $g / dm^3$): most acids involved with wine 
or fixed or nonvolatile (do not evaporate readily)

* **Volatile acidity** (acetic acid - $g / dm^3$): the amount of acetic acid in 
wine, which at too high of levels can lead to an unpleasant, vinegar taste

* **Citric acid** ($g / dm^3$): found in small quantities, citric acid can add 
'freshness' and flavor to wines

* **Residual sugar** ($g / dm^3$): the amount of sugar remaining after 
fermentation stops, it's rare to find wines with less than 1 $g/l$ and wines 
with greater than 45 $g/l$ are considered sweet

* **Chlorides** (sodium chloride - $g / dm^3$): the amount of salt in the wine

* **Free sulfur dioxide** ($mg / dm^3$): the free form of $SO_{2}$ exists in 
equilibrium between molecular $SO_{2}$ (as a dissolved gas) and bisulfite ion; 
it prevents microbial growth and the oxidation of wine

* **Total sulfur dioxide** ($mg / dm^3$): amount of free and bound forms of 
$SO_{2}$; in low concentrations, $SO_{2}$ is mostly undetectable in wine, but at
free $SO_{2}$ concentrations over 50 ppm, $SO_{2}$ becomes evident in the nose 
and taste of wine

* **Density** ($g / cm^3$): the density of water is close to that of water 
depending on the percent alcohol and sugar content

* **pH**: describes how acidic or basic a wine is on a scale from 0 (very 
acidic) to 14 (very basic); most wines are between 3-4 on the pH scale

* **Sulphates** (potassium sulphate - $g / dm3$): a wine additive which can 
contribute to sulfur dioxide gas ($SO_{2}$) levels, wich acts as an 
antimicrobial and antioxidant

* **Alcohol** (% volume): the percent alcohol content of the wine

* **Quality** (score between 0 and 10): output variable, based on sensory data


### Data Summary:

It will be interesting to combine white and red wine datasets as this could make 
richer the study. The process of merging them is quite direct as they present 
identical structure. In order to further analyses, we will add a categorical
variable called 'type' with 'red' and 'white' values.

The main scope of this study will be to analyze the main variables that affect 
the quality of the wine, for both red and white, and if they are the same. In 
order to simplify that task, we will add another variable called 'category' that
groups the existing 'quality' (1-10) into 3 new categories ('excellent', 
'average','deficient').

For the last sections of the study, we will also use a categorical variable for
the alcohol degree with 4 categories ('<9.5', '9.5-10.5', '10.5-11.5', '>11.5')

We will also have a look on some  similarities and differences between the 
types of wine throughout the analysis.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(ggplot2)
library(dplyr)
library(gridExtra)
library(corrplot)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE, include=FALSE}
# Load the Data
getwd()
setwd("C:/Users/Serg/Documents/Aprendiendo/Data Analysis/Module 5 - Exploratory Data Analysis (EDA)/Proyecto final")

# read the .csv files and create a dataframe for each kind of wine
red_wine <- read.csv('wineQualityReds.csv')
white_wine <- read.csv('wineQualityWhites.csv')

# add a column called "typè" to categorize each type of wine
red_wine['type'] <- 'red'
white_wine['type'] <- 'white'

# bind both datasets
wine_data <- rbind(red_wine, white_wine)

# add a column to group the quality into 3 categories
wine_data['category'] <- ifelse(wine_data$quality < 5, 'Deficient', ifelse(
  wine_data$quality < 7, 'Average', 'Excellent'))
wine_data$category<-factor(wine_data$category, levels=c("Deficient", "Average", 
                                                        "Excellent"))


# add a column to group the alcohol into 4 categories
wine_data['alcoholcat'] <- ifelse(wine_data$alcohol < 9.5, 'Alcohol < 9.5', ifelse(
  wine_data$alcohol < 10.5, 'Alcohol 9.5-10.5', ifelse(wine_data$alcohol < 11.5, 
                                               'Alcohol 10.5-11.5', 'Alcohol > 11,5')))
wine_data$alcoholcat<-factor(wine_data$alcoholcat, levels=c("Alcohol < 9.5", "Alcohol 9.5-10.5", 
                                                            "Alcohol 10.5-11.5", 
                                                            "Alcohol > 11,5"))

# remove X column
wine_data$X <- NULL
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
summary(wine_data)

```

We can appreciate that some variables have much variability (residual sugar, 
free and total sulfur dioxide, chlorides, acidity) due to outliers that can be
caused by bad measures for examples. We can also see some 0-values in the 
variable citric acid, which could be some kind of measuring error or lack of 
info of these datasets.


## Univariate Plots Section

In this section, we will have a look on the distribution of the different 
variables for both red and white wines and comparing their distributions to get 
a deeper understanding of the variables of interest and the key differences 
between both types of wine.

* **Quality histogram**:
```{r}
ggplot(wine_data, aes(quality)) +
  geom_histogram(binwidth = 1)+
    labs(x = 'Quality (score between 3 and 9)',
         y = 'Count')
```
 
 We can appreciaty that quality values are normally distributed for the wines.

* **Quality levels histogram**:

```{r}
ggplot(wine_data, aes(category)) +
  geom_bar()+
    labs(x = 'Quality Levels',
         y = 'Count')

```

As appreciated before, the quality (now grouped) is normally distributed.

* **Citric acid**:

```{r, echo=FALSE, message=FALSE, warning=FALSE}
p1_citric.acid <- ggplot(wine_data, aes(citric.acid)) + 
  geom_histogram() +
  facet_grid(type~.)+
  scale_x_continuous(lim = c(min(wine_data$citric.acid), 
                             quantile(wine_data$citric.acid,  
                                      probs = c(99.5)/100)))+
    labs(x = 'Citric Acid (g/dm3)',
         y = 'Count')
  
p2_citric.acid <- ggplot(wine_data, aes(type, citric.acid))+
  geom_boxplot(fill = 'lightblue')+ 
  stat_summary(fun.y=mean, color="orange", geom="point", 
               shape=18, size=3)+
  scale_y_continuous(lim = c(min(wine_data$citric.acid), 
                             quantile(wine_data$citric.acid,  
                                      probs = c(99.5)/100)))+
    labs(x = 'Wine Type',
         y = 'Citric Acid (g/dm3)')

grid.arrange(p1_citric.acid, p2_citric.acid, nrow = 1)
```


We can see a wide variability of citric acid for the red wine with a quite
homogeneous distribution, while white wine is normally distributed. In the
boxplot, we can appreciate that variability for the red wine, but both types 
don't have significantly different volumes of citric acid

* **Residual sugar**:

```{r, echo=FALSE, message=FALSE, warning=FALSE}
p1_residual.sugar <- ggplot(wine_data, aes(residual.sugar)) + 
  geom_histogram() +
  facet_grid(type~.)+
  scale_x_continuous(lim = c(min(wine_data$residual.sugar), 
                             quantile(wine_data$residual.sugar,  
                                      probs = c(99.5)/100)))+
    labs(x = 'Residual Sugar (g/dm3)',
         y = 'Count')
  
p2_residual.sugar <- ggplot(wine_data, aes(type, residual.sugar))+
  geom_boxplot(fill = 'lightblue')+ 
  stat_summary(fun.y=mean, color="orange", geom="point", 
               shape=18, size=3)+
  scale_y_continuous(lim = c(min(wine_data$residual.sugar), 
                             quantile(wine_data$residual.sugar,  
                                      probs = c(99.5)/100)))+
    labs(x = 'Wine Type',
         y = 'Residual Sugar (g/dm3)')

grid.arrange(p1_residual.sugar, p2_residual.sugar, nrow = 1)
```


In that case, we appreciate two right-skewed distribution and a clearly wider 
range of residual sugar volumes for the white wine (2-10) vs. red wine (~2). 
That makes sense because white wine tastes generally sweeter even though we can 
also see white wines with similar sugar volumes than white wine.

* **Chlorides**:

```{r, echo=FALSE, message=FALSE, warning=FALSE}
p1_chlorides <- ggplot(wine_data, aes(chlorides)) + 
  geom_histogram() +
  facet_grid(type~.)+
  scale_x_continuous(lim = c(min(wine_data$chlorides), 
                             quantile(wine_data$chlorides,  
                                      probs = c(99)/100)))+
    labs(x = 'Sodium Chloride (g/dm3)',
         y = 'Count')
  
p2_chlorides <- ggplot(wine_data, aes(type, chlorides))+
  geom_boxplot(fill = 'lightblue')+ 
  stat_summary(fun.y=mean, color="orange", geom="point", 
               shape=18, size=3)+
  scale_y_continuous(lim = c(min(wine_data$chlorides), 
                             quantile(wine_data$chlorides,  
                                      probs = c(99)/100)))+
    labs(x = 'Wine Type',
         y = 'Sodium Chloride (g/dm3)')

grid.arrange(p1_chlorides, p2_chlorides, nrow = 1)
```

Chlorides are normally distributed for red wine and right-skewed for white wine,
with a very long tail. We appreciate that white wine has significantly less
volume of chlorides than red wine.

* **Free sulfur dioxide**:

```{r, echo=FALSE, message=FALSE, warning=FALSE}
p1_free.sulfur.dioxide <- ggplot(wine_data, aes(free.sulfur.dioxide)) + 
  geom_histogram() +
  facet_grid(type~.)+
  scale_x_continuous(lim = c(min(wine_data$free.sulfur.dioxide), 
                             quantile(wine_data$free.sulfur.dioxide,  
                                      probs = c(99)/100)))+
    labs(x = 'Free Sulfur Dioxide (mg/dm3)',
         y = 'Count')

p2_free.sulfur.dioxide <- ggplot(wine_data, aes(type, free.sulfur.dioxide))+
  geom_boxplot(fill = 'lightblue')+ 
  stat_summary(fun.y=mean, color="orange", geom="point", 
               shape=18, size=3)+
  scale_y_continuous(lim = c(min(wine_data$free.sulfur.dioxide), 
                             quantile(wine_data$free.sulfur.dioxide,  
                                      probs = c(99)/100)))+
    labs(x = 'Wine Type',
         y = 'Free Sulfur Dioxide (mg/dm3)')

grid.arrange(p1_free.sulfur.dioxide, p2_free.sulfur.dioxide, nrow = 1)
```

Free sulfure dioxide's distribution is right-skewed for red wine and normal for
white wine. We can appreciate that white wine has significantly more 
concentration of free sulfure dioxide than red wine.

* **Total sulfur dioxide**:

```{r, echo=FALSE, message=FALSE, warning=FALSE}
p1_total.sulfur.dioxide <- ggplot(wine_data, aes(total.sulfur.dioxide)) + 
  geom_histogram() +
  facet_grid(type~.)+
  scale_x_continuous(lim = c(min(wine_data$total.sulfur.dioxide), 
                             quantile(wine_data$total.sulfur.dioxide,  
                                      probs = c(99)/100)))+
    labs(x = 'Total Sulfur Dioxide (mg/dm3)',
         y = 'Count')

p2_total.sulfur.dioxide <- ggplot(wine_data, aes(type, total.sulfur.dioxide))+
  geom_boxplot(fill = 'lightblue')+ 
  stat_summary(fun.y=mean, color="orange", geom="point", 
               shape=18, size=3)+
  scale_y_continuous(lim = c(min(wine_data$total.sulfur.dioxide), 
                             quantile(wine_data$total.sulfur.dioxide,  
                                      probs = c(99)/100)))+
    labs(x = 'Wine Type',
         y = 'Total Sulfur Dioxide (mg/dm3)')

grid.arrange(p1_total.sulfur.dioxide, p2_total.sulfur.dioxide, nrow = 1)
```

We appreciate a right-skewed distribution for red wine and a slightly 
left-skewed distribution for white wine. In that case, we can see that total
sulfure dioxide volumes are significantly higher for white wine.

* **Density**:

```{r, echo=FALSE, message=FALSE, warning=FALSE}
p1_density <- ggplot(wine_data, aes(density)) + 
  geom_histogram() +
  facet_grid(type~.)+
  scale_x_continuous(lim = c(min(wine_data$density), 
                             quantile(wine_data$density,  
                                      probs = c(99.9)/100)))+
    labs(x = 'Density (g/cm3)',
         y = 'Count')

p2_density <- ggplot(wine_data, aes(type, density))+
  geom_boxplot(fill = 'lightblue')+ 
  stat_summary(fun.y=mean, color="orange", geom="point", 
               shape=18, size=3)+
  scale_y_continuous(lim = c(min(wine_data$density), 
                             quantile(wine_data$density,  
                                      probs = c(99.9)/100)))+
    labs(x = 'Wine Type',
         y = 'Density (g/dm3)')

grid.arrange(p1_density, p2_density, nrow = 1)
```

Both wines seem to be normally distributed and red wines seem to have slightly
higher density than white even though their boxplots overlap.

* **pH**:

```{r, echo=FALSE, message=FALSE, warning=FALSE}
p1_pH <- ggplot(wine_data, aes(pH)) + 
  geom_histogram() +
  facet_grid(type~.)+
  scale_x_continuous(lim = c(min(wine_data$pH), 
                             quantile(wine_data$pH,  probs = c(99.9)/100)))+
    labs(x = 'pH (0-14)',
         y = 'Count')

p2_pH <- ggplot(wine_data, aes(type, pH))+
  geom_boxplot(fill = 'lightblue')+ 
  stat_summary(fun.y=mean, color="orange", geom="point", 
               shape=18, size=3)+
  scale_y_continuous(lim = c(min(wine_data$pH), 
                             quantile(wine_data$pH,  probs = c(99.9)/100)))+
    labs(x = 'Wine Type',
         y = 'pH (0-14)')

grid.arrange(p1_pH, p2_pH, nrow = 1)
```

pH is normally distributed for both wines and seems slightly higher for red 
wine.

* **Sulphates**:

```{r, echo=FALSE, message=FALSE, warning=FALSE}
p1_sulphates <- ggplot(wine_data, aes(sulphates)) + 
  geom_histogram() +
  facet_grid(type~.)+
  scale_x_continuous(lim = c(min(wine_data$sulphates), 
                             quantile(wine_data$sulphates,  
                                      probs = c(99.5)/100)))+
    labs(x = 'Potassium Sulphate (g/dm3)',
         y = 'Count')

p2_sulphates <- ggplot(wine_data, aes(type, sulphates))+
  geom_boxplot(fill = 'lightblue')+ 
  stat_summary(fun.y=mean, color="orange", geom="point", 
               shape=18, size=3)+
  scale_y_continuous(lim = c(min(wine_data$sulphates), 
                             quantile(wine_data$sulphates,  
                                      probs = c(99.5)/100)))+
    labs(x = 'Wine Type',
         y = 'Potassium Sulphate (g/dm3)')

grid.arrange(p1_sulphates, p2_sulphates, nrow = 1)
```

Sulphates have a right-skewed distribution for both wines. Its concentration
seems to be slightly higher for red wine.

* **Alcohol**:

```{r, echo=FALSE, message=FALSE, warning=FALSE}
p1_alcohol <- ggplot(wine_data, aes(alcohol)) + 
  geom_histogram() +
  facet_grid(type~.)+
  scale_x_continuous(lim = c(min(wine_data$alcohol), 
                             quantile(wine_data$alcohol,  
                                      probs = c(99.9)/100)))+
    labs(x = 'Alcohol (% volume)',
         y = 'Count')

p2_alcohol <- ggplot(wine_data, aes(type, alcohol))+
  geom_boxplot(fill = 'lightblue')+ 
  stat_summary(fun.y=mean, color="orange", geom="point", 
               shape=18, size=3)+
  scale_y_continuous(lim = c(min(wine_data$alcohol), 
                             quantile(wine_data$alcohol,  
                                      probs = c(99.9)/100)))+
    labs(x = 'Wine Type',
         y = 'Alcohol (% volume)')

grid.arrange(p1_alcohol, p2_alcohol, nrow = 1)
```

Alcohol volume distribution is right-skewed for both wines. We can't appreciate
that both types of wine have significantly different alcohol degrees as they
have very similar boxplots and mean.


## Univariate Analysis

#### What is the structure of your dataset?
We have a sample size of 6497 wines (1599 red and 4898 white).

#### What is/are the main feature(s) of interest in your dataset?
The main feature will be quality and we will try to determine the variables that
make wine taste better.

#### What other features in the dataset do you think will help support your 
#### investigation into your feature(s) of interest?
Probably, alcohol, residual sugar, pH, acid and density will affect the quality.

#### Did you create any new variables from existing variables in the dataset?
Yes, I added one column for the type of wine (red or white) and another one for
the quality divided in three groups (excellent, average and deficient)

#### Of the features you investigated, were there any unusual distributions? 
#### Did you perform any operations on the data to tidy, adjust, or change the 
#### form of the data? If so, why did you do this?
There where outliers in most of the features. I trimmed some outliers little
quantiles (<0.1) when plotting their distributions in order to appreciate them 
clearer.

## Bivariate Plots Section

* **Red&White feature correlation together:** 

```{r, echo=FALSE, fig.height=10, fig.width=10, message=FALSE, warning=FALSE}
wine_data$numcategory <- as.numeric(wine_data$quality)
mtx <- cor(wine_data[c('fixed.acidity', 'volatile.acidity', 'citric.acid', 
                       'residual.sugar', 'chlorides', 'free.sulfur.dioxide', 
                       'total.sulfur.dioxide', 'density','pH', 'sulphates', 
                       'alcohol', 'numcategory')])

wine_corrplot <- corrplot(mtx, type="upper", 
                          order="hclust", 
                          tl.col="black", 
                          tl.srt=45)

mtx
```

For the merged dataset, we can easily appreciate on the plot the strong inverse
correlation between alcohol and density (-0.69), what makes sense because a 
higher alcohol degree is lowering the density of the wine. We also see other
pairs of variables strongly correlated like total.sulfur.dioxide and 
free.sulfure.dioxide (0.72) and some others with correlation >=0.5: total sulfur
dioxide and residual
sugar, density and residual sugar.

* **Red Wine feature correlation:**

```{r, echo=FALSE, fig.height=10, fig.width=10, message=FALSE, warning=FALSE}
mtx_red <- cor(subset(wine_data, 
                      wine_data['type'] == 'red')[c('fixed.acidity',
                                                    'volatile.acidity', 
                                                    'citric.acid',
                                                    'residual.sugar',
                                                    'chlorides', 
                                                    'free.sulfur.dioxide',
                                                    'total.sulfur.dioxide', 
                                                    'density','pH', 'sulphates',
                                                    'alcohol', 'numcategory')])
redwine_corrplot <- corrplot(mtx_red, type="upper", 
                             order="hclust", 
                             tl.col="black", 
                             tl.srt=45)

mtx_red
```

For the red wine data, we can see strong inverse correlation between alcohol
fixed acidity and pH (-0.68), citric acid and pH (-0.54), and citric acid and 
volatile acidity (-0.55). We also find direct correlation between fixed acidity 
and citric acid (0.67), density and fixed acidity (0.67), and total sulfur 
dioxide and free sulfur dioxide (0.67).

Regarding the variable of interest, it gets its higher correlation with alcohol 
(0.48) for this wine.

* **White Wine feature correlation:**

```{r, echo=FALSE, fig.height=10, fig.width=10, message=FALSE, warning=FALSE}
mtx_white <- cor(subset(wine_data, wine_data['type'] == 'white')[c('fixed.acidity',
                                                    'volatile.acidity', 
                                                    'citric.acid',
                                                    'residual.sugar',
                                                    'chlorides', 
                                                    'free.sulfur.dioxide',
                                                    'total.sulfur.dioxide', 
                                                    'density','pH', 'sulphates',
                                                    'alcohol', 'numcategory')])

whitewine_corrplot <- corrplot(mtx_white, type="upper", 
                               order="hclust", 
                               tl.col="black", 
                               tl.srt=45)

mtx_white
```

For the white wine data, we can see strong inverse correlation between density
and alcohol (-0.78) and density and residual sugar (-0.84). Also, total sulfur 
dioxide is significantly correlated with free sulfur dioxide (0.62) and density 
(0.53). In this case, there are not any other correlations equal or greater than
0.5.

Regarding the variable of interest, it gets its higher correlation with alcohol 
(0.42) for this wine.

* **Density vs. Alcohol:**

```{r, echo=FALSE, message=FALSE, warning=FALSE}

ggplot(wine_data, aes(density, alcohol)) +
  facet_wrap(~type, ncol = 1) +  
  geom_point(alpha = 1/4) + 
  stat_smooth(method = 'lm') +
  xlim(min(wine_data$density), quantile(wine_data$density, 0.95))+
    labs(x = 'Density (g/cm3)',
         y = 'Alcohol (% volume)') 


```

We can appreciate the better lineal fit for the white wine (-0.78) compared to
the red wine (-0.45).

* **Density vs. Residual sugar:**

```{r, echo=FALSE, message=FALSE, warning=FALSE}

ggplot(wine_data, aes(density, residual.sugar)) +
  facet_wrap(~type, ncol = 1) +  
  geom_point(alpha = 1/4) + 
  stat_smooth(method = 'lm') +
  xlim(min(wine_data$density), quantile(wine_data$density, 0.95)) + 
  ylim(min(wine_data$residual.sugar), quantile(wine_data$residual.sugar, 0.95))+
    labs(x = 'Density (g/cm3)',
         y = 'Residual Sugar (g/dm3)') 

```

We can also see a strong correlation between residual sugar and density in the
white wine (0.84), while red wine has a notably lower correlation (0.36). This 
makes sense as white wine use to be sweeter and that quantity of residual sugar
makes the difference for the density.

* **Total sulfur dioxide vs. Free sulfur dioxide:**

```{r, echo=FALSE, message=FALSE, warning=FALSE}

ggplot(wine_data, aes(total.sulfur.dioxide, free.sulfur.dioxide)) +
  facet_wrap(~type, ncol = 1) +  
  geom_point(alpha = 1/4) + 
  stat_smooth(method = 'lm') +
  xlim(min(wine_data$total.sulfur.dioxide), 
       quantile(wine_data$total.sulfur.dioxide, 0.95)) + 
  ylim(min(wine_data$free.sulfur.dioxide), 
       quantile(wine_data$free.sulfur.dioxide, 0.95))+
    labs(x = 'Total Sulfur Dioxide (mg/dm3)',
         y = 'Free Sulfur Dioxide (mg/dm3)') 

```

In this case, both have similar correlations: red wine (0.67) and white wine 
(0.62).

* **Quality vs. Alcohol:**

```{r, echo=FALSE, message=FALSE, warning=FALSE}

ggplot(wine_data, aes(quality, alcohol)) +
  facet_wrap(~type, ncol = 1) +  
  geom_jitter(alpha = 1/4) + 
  stat_smooth(method = 'lm') +
  ylim(min(wine_data$alcohol), quantile(wine_data$alcohol, 0.95))+
    labs(x = 'Quality (score between 3 and 9)',
         y = 'Alcohol (% volume)') 

```

The correlation for both wines exist, even though is not the strongest (0.4-0.5)

__________________________________

Based on what we have just found out on correlations, we will break down the 
quality in three categories and analyze what features affect the most to good 
and bad wines.

* **Category vs. Alcohol:**


```{r, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(wine_data, aes(category, alcohol)) +
  geom_boxplot()+ 
  facet_wrap(~type, ncol = 2) +  
  stat_summary(fun.y=mean, color="orange", geom="point", 
               shape=18, size=3)+
    labs(x = 'Quality Levels',
         y = 'Alcohol (% volume)')  
```

More alcohol, usually means a significative better wine for graduation higher 
than 11 degrees.

* **Category vs. Density:**


```{r, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(wine_data, aes(category, density)) +
  geom_boxplot()+ 
  facet_wrap(~type, ncol = 2) +  
  coord_cartesian(ylim = c(0.987, 1.004)) +
  stat_summary(fun.y=mean, color="orange", geom="point", 
               shape=18, size=3)+
    labs(x = 'Quality Levels',
         y = 'Density (g/cm3)') 
```

For white wines, the excellent wines seem to have a lower density, what makes 
sense as density has a strong inverse correlation with the degree of alcohol.

* **Category vs. Volatile Acidity:**


```{r, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(wine_data, aes(category, volatile.acidity)) +
  geom_boxplot()+ 
  facet_wrap(~type, ncol = 2) +  
  coord_cartesian(ylim = c(0, 1.25)) +
  stat_summary(fun.y=mean, color="orange", geom="point", 
               shape=18, size=3)+
    labs(x = 'Quality Levels',
         y = 'Acetic Acid (g/dm3)') 
```

For red wines, higher concentrations of acetic acid (volatile acidity) mean 
worse wines. The lower concentration, the better seems to be the red wine.


## Bivariate Analysis

#### Talk about some of the relationships you observed in this part of the
#### investigation. How did the feature(s) of interest vary with other features 
#### in the dataset?

The feature of interest has an important direct correlation with alcohol. It 
means that a bigger degree of alcohol means a higher quality of the wine generally.

Breaking down the quality in three categories, we find out that the excellent 
wines for both red and white types seem to be those with an alcohol degree 
greater than 11.

#### Did you observe any interesting relationships between the other features 
#### (not the main feature(s) of interest)?

Yes, there are some interesting relationships between density and alcohol and
density and residual sugar, that makes us think of the rol of sugar in white 
wine and the lower density of alcohol compared to water, for example.

#### What was the strongest relationship you found?

It was density vs. residual sugar for the white wine: a direct correlation of 
0.84.



## Multivariate Plots Section

For the following plots, we defined another categorical variable that splits up 
the degree of alcohol in 4 categories, as we saw the importance of this feature
in the quality of both red and white wine

* **Red wine:Alcohol, Quality, Citric Acid and Acetic Acid**

```{r, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(aes(x = citric.acid, y = volatile.acidity, color = category), 
       data = subset(wine_data, wine_data$type == 'red')) +
    facet_wrap(~alcoholcat, ncol =4) + 
    geom_jitter(size = 2, alpha = 0.5) +
    scale_colour_brewer(palette = 'Blues', guide =guide_legend('Quality Levels'))+
    xlim(min(wine_data$citric.acid), quantile(wine_data$citric.acid, 0.95)) +
    ylim(min(wine_data$volatile.acidity), 
         quantile(wine_data$volatile.acidity, 0.95))+
    labs(x = 'Citric Acid (g/dm3)',
         y = 'Acetic Acid (g/dm3)')
```
 
 In this plot, we can clearly observe a trend for red wine by which the wines 
 with higher level of alcohol, less concentration of acetic acid and more
 concentration of citric acid, get the better qualifications.

* **White wine: Alcohol, Quality, Chlorides and Density**


```{r, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(aes(x = density, y = chlorides, color = category), 
       data = subset(wine_data, wine_data$type == 'white')) +
    facet_wrap(~alcoholcat, ncol =4) + 
    geom_point(size = 1, alpha = 0.5) +
    scale_colour_brewer(palette = 'Blues', guide =guide_legend('Quality Levels'))+
    xlim(min(wine_data$density), quantile(wine_data$density, 0.95)) +
    ylim(min(wine_data$chlorides), quantile(wine_data$chlorides, 0.95))+
    labs(x = 'Density (g/cm3)',
         y = 'Sodium Chloride (g/dm3)')
```

For white wines, the best wines have a higher degree of alcohol, a lower 
concentration of chlorides and a lower density (correlated to a high level of
residual sugar as well)

## Multivariate Analysis

#### Talk about some of the relationships you observed in this part of the 
#### investigation. Were there features that strengthened each other in terms of
#### looking at your feature(s) of interest?

We pointed out that alcohol was the main feature affecting the quality of the
wine. However, we added 2 more features to the equation, the 2 features with 
higher correlation with quality after the alcohol degree.

Doing this allow us to visualize clearly, for red and white wine, that these 
features shift the scatter plot and make the difference.

#### Were there any interesting or surprising interactions between features?

Yes, citric and acetic acid concentration for red wine and density and chlorides 
for white wine.

------

## Final Plots and Summary

#### Plot One: Degree of alcohol. Difference between red and white?

```{r, echo=FALSE, message=FALSE, warning=FALSE}
p1_alcohol <- ggplot(wine_data, aes(alcohol)) + 
  geom_histogram() +
  facet_grid(type~.)+
  scale_x_continuous(lim = c(min(wine_data$alcohol), 
                             quantile(wine_data$alcohol,  
                                      probs = c(99.9)/100)))+
    labs(x = 'Alcohol (% volume)',
         y = 'Count')

p2_alcohol <- ggplot(wine_data, aes(type, alcohol))+
  geom_boxplot(fill = 'lightblue')+ 
  stat_summary(fun.y=mean, color="orange", geom="point", 
               shape=18, size=3)+
  scale_y_continuous(lim = c(min(wine_data$alcohol), 
                             quantile(wine_data$alcohol,  
                                      probs = c(99.9)/100)))+
    labs(x = 'Wine Type',
         y = 'Alcohol (% volume)')

grid.arrange(p1_alcohol, p2_alcohol, nrow = 1)
```


#### Description One
Personally, I had always thought that white wines had significantly more degree
of alcohol but in these plots we can clearly see very similar distributions, as 
well as parameters (mean, median and quantiles). So that, the alcohol doesn't
seem to be a significative feature between both types of wine.


#### Plot Two: Boxplot: wine category and alcohol degree
```{r, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(wine_data, aes(category, alcohol)) +
  geom_boxplot()+ 
  facet_wrap(~type, ncol = 2) +  
  stat_summary(fun.y=mean, color="orange", geom="point", 
               shape=18, size=3)+
    labs(x = 'Quality Levels',
         y = 'Alcohol (% volume)')  
```


#### Description Two
We observed that the highest correlation between one feature and the quality of 
the wine was the degree of alcohol for both wines, even though it is not a 
strong correlation: 0.48 for red wine and 0.43 for white wine.

With this plot two, we break down the samples of both types of wine in 
deficient, average and excellent wines. And, surprisingly, excellent wines seem to be a
higher degree of alcohol (more than 11 degrees). It seemed misleading for me 
because drinks with higher degree of alcohol usually taste stronger.


#### Plot Three: Alcohol, Quality, Chlorides and Density

```{r, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(aes(x = density, y = chlorides, color = category), 
       data = subset(wine_data, wine_data$type == 'white')) +
    facet_wrap(~alcoholcat, ncol =4) + 
    geom_point(size = 1, alpha = 0.5) +
    scale_colour_brewer(palette = 'Blues', guide =guide_legend('Quality Levels'))+
    xlim(min(wine_data$density), quantile(wine_data$density, 0.95)) +
    ylim(min(wine_data$chlorides), quantile(wine_data$chlorides, 0.95))+
    labs(x = 'Density (g/cm3)',
         y = 'Sodium Chloride (g/dm3)')
```




#### Description Three
For white wines, the best wines have a higher degree of alcohol, a lower 
concentration of chlorides and a lower density (correlated to a high level of
residual sugar as well)


------

## Reflection

By analyzing this data, we have come to some ideas that can be new for some of 
us.

1. Red and white wine are different in many aspects (citric acid, residual 
sugar, chlorides, sulfure dioxide, density, pH,...) but the one they are more 
similar is the one which has more to do with the taste of excellent wines: the 
alcohol degree

2. Nevertheless, we observe different features for each kind of wine that affect
the flavor: citric acid and acetic acid concentration for red wine and chlorides 
and density (affected as well for the concentration of residual sugar) for white
wine

So according to this analysis, the main feature that affects the flavor of a 
good wine is the degree of alcohol, which should be higher than 11. Moreover,
if you are going to produce your own wine, you should take into account the 
other features we commented for red and white wine, so you can assure an 
excellent taste for your production.

I got struggle with merging the data at first and with plotting correctly
because of the different scales. I had to visit often "Stackoverflow" in order
to get through course limitation and my own.

Next steps of this analysis may include a larger number of samples for the red
wine dataset so we have similar number of samples in both datasets that we
merged. We can also consider other kind of plots and add a linear model.

If we got a large enough dataset, we could also create a predictive model with
Machine Learning algorithms and use it to identify the color of the wine or the
expected taste based on the other parameters.

For future exploration of this data I would pick one category of wine (for 
example, quality level 3-4, 5-7, 8-9) to look at the patterns which can appear 
in each of these three buckets. I also would normalize data because we have more
white wine than red wine.