---
title: "diamonds (lesson 4)"
---
## Load diamonds dataset, included in the ggplot2 library
```{r}
library(ggplot2)
data(diamonds)
summary(diamonds)
```

## Let's have a look on the explanation of diamonds dataset
```{r}
?diamonds
```

# Create a histogram of the price of
# all the diamonds in the diamond data set.

# TYPE YOUR CODE BELOW THE LINE
# =======================================
```{r}
qplot(data = diamonds, x = price, binwidth = 500)
summary(diamonds$price)
```

## How many diamonds cost less than $500?
```{r}
nrow(subset(diamonds, price < 500))
```

## How many diamonds cost less than $250?
```{r}
nrow(subset(diamonds, price < 250))
```

## How many diamonds cost $15,000 or more?
```{r}
nrow(subset(diamonds, price >= 15000))
```

# Explore the largest peak in the
# price histogram you created earlier.

# Try limiting the x-axis, altering the bin width,
# and setting different breaks on the x-axis.

# There won’t be a solution video for this
# question so go to the discussions to
# share your thoughts and discover
# what other people find.

# You can save images by using the ggsave() command.
# ggsave() will save the last plot created.
# For example...
#                  qplot(x = price, data = diamonds)
#                  ggsave('priceHistogram.png')

# ggsave currently recognises the extensions eps/ps, tex (pictex),
# pdf, jpeg, tiff, png, bmp, svg and wmf (windows only).

# Submit your final code when you are ready.

# TYPE YOUR CODE BELOW THE LINE
# ======================================================================
```{r}
qplot(data = diamonds, x = price, binwidth = 25) +
  scale_x_continuous(lim = c(326, 2400))
```

# Break out the histogram of diamond prices by cut.

# You should have five histograms in separate
# panels on your resulting plot.

# TYPE YOUR CODE BELOW THE LINE
# ======================================================
```{r}
qplot(data = diamonds, x = price) + 
  facet_wrap(~cut, scales = "free_y")

by(diamonds$price, diamonds$cut, summary)

```

# Create a histogram of price per carat
# and facet it by cut. You can make adjustments
# to the code from the previous exercise to get
# started.

# Adjust the bin width and transform the scale
# of the x-axis using log10.

# Submit your final code when you are ready.

# ENTER YOUR CODE BELOW THIS LINE.
# ===========================================================================
```{r}
ggplot(diamonds) + geom_histogram(aes(x = price/carat), binwidth = 0.05,
                                  color = "black", fill = "orange") +
  scale_x_log10() +
  facet_grid(~cut)
```

# Investigate the price of diamonds using box plots,
# numerical summaries, and one of the following categorical
# variables: cut, clarity, or color.

# There won't be a solution video for this
# exercise so go to the discussion thread for either
# BOXPLOTS BY CLARITY, BOXPLOT BY COLOR, or BOXPLOTS BY CUT
# to share you thoughts and to
# see what other people found.

# You can save images by using the ggsave() command.
# ggsave() will save the last plot created.
# For example...
#                  qplot(x = price, data = diamonds)
#                  ggsave('priceHistogram.png')

# ggsave currently recognises the extensions eps/ps, tex (pictex),
# pdf, jpeg, tiff, png, bmp, svg and wmf (windows only).

# Copy and paste all of the code that you used for
# your investigation, and submit it when you are ready.
# =================================================================

```{r}
by(diamonds$price, diamonds$color, summary)
qplot(x = color, y = price, data = diamonds,
      geom = 'boxplot') + coord_cartesian(ylim = c(0, 8000))
ggplot(diamonds, aes(y = price)) + geom_boxplot(aes(x = color))
IQR(subset(diamonds, color == "J")$price)
IQR(subset(diamonds, color == "D")$price)

```

```{r}
by(diamonds$price, diamonds$cut, summary)
qplot(x = cut, y = price, data = diamonds,
      geom = 'boxplot') + coord_cartesian(ylim = c(0, 8000))
ggplot(diamonds, aes(y = price)) + geom_boxplot(aes(x = cut))

```




```{r}
by(diamonds$price, diamonds$clarity, summary)
qplot(x = clarity, y = price, data = diamonds,
      geom = 'boxplot') + coord_cartesian(ylim = c(0, 8000))
ggplot(diamonds, aes(y = price)) + geom_boxplot(aes(x = clarity))

```

```{r}
ggplot(diamonds, aes(y = price/carat)) + geom_boxplot(aes(x = color)) + 
    ylab("price / carat") + coord_cartesian(ylim = c(2000, 6000))
```

```{r}
ggplot(diamonds, aes(y = price/carat)) + geom_boxplot(aes(x = cut)) + 
    ylab("price / carat") + coord_cartesian(ylim = c(2000, 6000))
```

```{r}
ggplot(diamonds, aes(y = price/carat)) + geom_boxplot(aes(x = clarity)) + 
    ylab("price / carat") + coord_cartesian(ylim = c(2000, 6000))
```

```{r}
ggplot(diamonds) + geom_freqpoly(aes(x = carat), binwidth = 0.01,
                                  color = "red") +
                  scale_x_continuous(breaks = seq(0, 5, 0.1))
```


