Lesson 4
========================================================

***

### Scatterplots and Perceived Audience Size
Notes:
"People guess 20,50, when the actual number is more like 100 or 200"

***

### Scatterplots
Notes:

```{r Scatterplots}
library(ggplot2)
pf <- read.csv("pseudo_facebook.tsv", sep = "\t")

```

```{r}

qplot(x= age, y = friend_count, data = pf)

qplot(age, friend_count, data = pf)

```

***

#### What are some things that you notice right away?
Response:

***

### ggplot Syntax
Notes:

```{r ggplot Syntax}

ggplot(aes(x = age, y = friend_count), data = pf) +
  geom_point() +
  xlim(13, 90)

ggplot(aes(age, friend_count), data = pf) +
  geom_point()+
  xlim(13, 90)

```
```{r}
summary(pf$age)

```

***

### Overplotting
Notes:

```{r Overplotting}
ggplot(aes(x = age, y = friend_count), data = pf) +
  geom_point(alpha= 1/20) +
  xlim(13, 90)

ggplot(aes(age, friend_count), data = pf) +
  geom_point(alpha = 1/20)+
  xlim(13, 90)
```
```{r}
ggplot(aes(x = age, y = friend_count), data = pf) +
  geom_jitter(alpha= 1/20) +
  xlim(13, 90)

ggplot(aes(age, friend_count), data = pf) +
  geom_jitter(alpha = 1/20)+
  xlim(13, 90)
```

#### What do you notice in the plot?
Response:
Now, there isn't that clear the limit of each column. I have also realised that the number of removed rows have increased in 276
***

### Coord_trans()
Notes:
```{r}
?coord_trans()
```

#### Look up the documentation for coord_trans() and add a layer to the plot that transforms friend_count using the square root function. Create your plot!

```{r}
ggplot(aes(age, friend_count), data = pf) +
  geom_point(alpha = 1/20)+
  xlim(13, 90) +
  coord_trans(y = "sqrt")
```

#### What do you notice?

***

### Alpha and Jitter
Notes:

```{r Alpha and Jitter}
ggplot(aes(age, friendships_initiated), data = pf) +
  geom_jitter(alpha = 1/10, position = position_jitter(height = 0)) + 
  coord_trans(y = 'sqrt') + 
  xlim(13,90)
```
```{r}
summary(pf)
```

***

### Overplotting and Domain Knowledge
Notes:

***

### Conditional Means
Notes:

```{r Conditional Means}
install.packages('dplyr')
library(dplyr)
```



```{r Conditional Means}
age_groups <- group_by(pf, age)
pf.fc_by_age <- summarise(age_groups,
                          friend_count_mean = mean(friend_count),
                          friend_count_median = median(friend_count),
                          n = n()) 
pf.fc_by_age <- arrange(pf.fc_by_age, age)
head(pf.fc_by_age)
```
```{r Conditional Alternative Means Plot}
pf.fc_by_age <- pf %>%
  group_by(age) %>%
  summarise(friend_count_mean = mean(friend_count),
            friend_count_median = median(friend_count),
            n = n()) %>%
  arrange(age)
head(pf.fc_by_age, 20)
```

Create your plot!
```{r Conditional means plot}
ggplot(pf.fc_by_age, aes(age, friend_count_mean)) + 
  geom_line() +
  geom_point(alpha = 0.2) +
  xlim(13, 90) 
```

***

### Overlaying Summaries with Raw Data
Notes:

```{r Overlaying Summaries with Raw Data}
ggplot(pf,aes(age, friend_count)) +
  xlim(13, 90) +
  geom_point(alpha = 0.05, 
             position = position_jitter(h = 0),
             color = 'orange') +
  coord_trans(y = 'sqrt') +
  geom_line(stat = 'summary', fun.y = mean) +
  geom_line(stat = 'summary', fun.y = quantile, fun.args = list(probs = .1), linetype = 2,  color = 'blue') +
  geom_line(stat = 'summary', fun.y = quantile, fun.args = list(probs = .5), color = 'blue') +
  geom_line(stat = 'summary', fun.y = quantile, fun.args = list(probs = .9), linetype = 2,  color = 'blue')     
```

#### What are some of your observations of the plot?
Response:
Less than 10% of people have 1000+ friends and for people older than 33-35 it is common to be below 250 friends
***
```{r}
ggplot(pf,aes(age, friend_count)) +
  coord_cartesian(xlim = c(13, 70), ylim = c(0, 1e3)) + 
  geom_point(alpha = 0.05, 
             position = position_jitter(h = 0),
             color = 'orange') +
  geom_line(stat = 'summary', fun.y = mean) +
  geom_line(stat = 'summary', fun.y = quantile, fun.args = list(probs = .1), linetype = 2,  color = 'blue') +
  geom_line(stat = 'summary', fun.y = quantile, fun.args = list(probs = .5), color = 'blue') +
  geom_line(stat = 'summary', fun.y = quantile, fun.args = list(probs = .9), linetype = 2,  color = 'blue')
```

### Moira: Histogram Summary and Scatterplot
See the Instructor Notes of this video to download Moira's paper on perceived audience size and to see the final plot.

Notes:

***

### Correlation
Notes:

```{r Correlation}
?cor.test
cor.test(pf$age, pf$friend_count, method = 'pearson')

with(pf, cor.test(age, friend_count, method = 'pearson'))

```

Look up the documentation for the cor.test function.

What's the correlation between age and friend count? Round to three decimal places.
Response:

***

### Correlation on Subsets
Notes:
Cor.test uses Pearson method by default!

```{r Correlation on Subsets}
with(       subset(pf, age <= 70 )          , cor.test(age, friend_count))
```

***

### Correlation Methods
Notes:




```{r}
with(       subset(pf, age <= 70 )          , cor.test(age, friend_count,
                                                       method = 'spearman'))

```
***

## Create Scatterplots
Notes:
```{r}
ggplot(pf, aes(www_likes_received, likes_received)) +
  geom_point() +
  xlim(0, 1e4) + 
  ylim(0, 2e4) 
  
```

***

### Strong Correlations
Notes:

```{r Strong Correlations}
ggplot(pf, aes(www_likes_received, likes_received)) +
  geom_point() +
  xlim(0, quantile(pf$www_likes_received, 0.95)) + 
  ylim(0, quantile(pf$likes_received, 0.95)) +
  geom_smooth(method = 'lm', color = 'red')


```

What's the correlation betwen the two variables? Include the top 5% of values for the variable in the calculation and round to 3 decimal places.

```{r Correlation Calcuation}
with(pf, cor.test(www_likes_received, likes_received))
```

Response:

***

### Moira on Correlation
Notes:

***

### More Caution with Correlation
Notes:

```{r More Caution With Correlation}
install.packages('alr3')
library(alr3)
data(Mitchell)
```
```{r}
?Mitchell
summary(Mitchell)
```

Create your plot!

```{r Temp vs Month}
ggplot(Mitchell, aes(Month, Temp)) + 
  geom_point()
```

***

### Noisy Scatterplots
a. Take a guess for the correlation coefficient for the scatterplot.
0.8
b. What is the actual correlation of the two variables?
(Round to the thousandths place)

```{r Noisy Scatterplots}
with(Mitchell, cor.test(Month, Temp))
cor.test(Mitchell$Month, Mitchell$Temp)
```

***

### Making Sense of Data
Notes:
O
```{r Making Sense of Data}
ggplot(Mitchell, aes(Month, Temp)) + 
  geom_point() +
  scale_x_continuous(breaks = seq(0, 203, 12))
```
```{r}
range(Mitchell$Month)
```

***

### A New Perspective

What do you notice?
Response:

Watch the solution video and check out the Instructor Notes!
Notes:
```{r}
ggplot(Mitchell, aes(Month%%12, Temp))+
  geom_point()
```

***

### Understanding Noise: Age to Age Months
Notes:

```{r Understanding Noise: Age to Age Months}
ggplot(pf.fc_by_age, aes(age, friend_count_mean)) + 
  geom_line()

head(pf.fc_by_age, 10)
pf.fc_by_age[17:19, ]
```

***

### Age with Months Means

```{r Age with Months Means}
pf$age_with_months <- with(pf, age + (1 - dob_month / 12))


```
```{r}
?group_by
head(age_groups,100)
head(pf)
```

Programming Assignment
```{r Programming Assignment}
library(dplyr)
pf.fc_by_age_month <- pf %>%
  group_by(age_with_months) %>%
  summarise(friend_count_mean = mean(friend_count),
            friend_count_median = median(friend_count),
            n = n())%>%
  arrange(age_with_months)
```

```{r Alternative solution}
age_with_months_groups <- group_by(pf, age_with_months)
pf.fc_by_age_month2 <- summarise(age_with_months_groups,
                                 friend_count_mean = mean(friend_count),
                                 friend_count_median = median(friend_count),
                                 n = n())
pf.fc_by_age_month2 <- arrange(pf.fc_by_age_month2, age_with_months)
```

***

### Noise in Conditional Means

```{r Noise in Conditional Means}
ggplot(subset(pf.fc_by_age_month, age_with_months < 71), aes(age_with_months, friend_count_mean)) + 
  geom_line()

?subset
```

***

### Smoothing Conditional Means
Notes:

```{r Smoothing Conditional Means}
p1 <- ggplot(subset(pf.fc_by_age, age < 71), aes(age, friend_count_mean)) + 
  geom_line() + 
  geom_smooth()

p2 <- ggplot(subset(pf.fc_by_age_month, age_with_months < 71), aes(age_with_months, friend_count_mean)) + 
  geom_line() +
  geom_smooth()

p3 <- ggplot(subset(pf, age < 71), aes(round(age/5)*5, friend_count)) +
  geom_line(stat = 'summary', fun.y = mean)
library(gridExtra)
grid.arrange(p2, p1, p3, ncol = 1)
```

***

### Which Plot to Choose?
Notes:

***

### Analyzing Two Variables
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

