---
title: "Price vs X"
output: html_notebook
---
# Problem Set: Explore Two Variables

```{r}
# In this problem set, you'll continue
# to explore the diamonds data set.

# Your first task is to create a
# scatterplot of price vs x.
# using the ggplot syntax.

# Import Library
library(ggplot2)

# Load the dataset
data("diamonds")

# Check the Docs on the dataset
# ?diamonds

# Get the first few records
head(diamonds)
```
## Price vs. x

```{r}
# Inspiration

# Create scatterplot - using QPlot
# qplot(x = age, y = friend_count, data = pf)
#
# Create scatterplot - using GGPlot
# ggplot(aes(x = age, y = friend_count), data = pf)  + geom_point()

# Create scatterplot - using GGPlot
ggplot(diamonds, aes(depth, price))  + geom_point()
```
## Correlations

```{r}
# Correlation between Price and X,y and z
cor.test(diamonds$price, diamonds$x, method = 'pearson')
cor.test(diamonds$price, diamonds$y, method = 'pearson')
cor.test(diamonds$price, diamonds$z, method = 'pearson')

```
## Price vs. depth

```{r}
# Create a simple scatter plot of price vs depth.

# Create scatterplot - using GGPlot
ggplot(diamonds, aes(depth, price)) +
  geom_point()
```
## Adjustments - Price vs. depth

```{r}
# Change the code to make the transparency of the
# points to be 1/100 of what they are now and mark
# the x-axis every 2 units. See the instructor notes
# for two hints.

# Inspiration:
#  
# Jitter
# ggplot(aes(x = age, y = friend_count), data = pf)  + geom_jitter(alpha = 1/20) + xlim(13, 90)

# Create scatterplot - using GGPlot, change the transparancy
ggplot(diamonds, aes(depth, price)) + 
  geom_point(alpha = 1/100) +
  scale_x_continuous(breaks = seq(0, 100, 2))
```
## Correlation - price and depth
```{r}
with(diamonds, cor.test(price, depth, method = 'pearson'))
```
## Price vs. carat
```{r}
# Create a scatterplot of price vs carat
# and omit the top 1% of price and carat
# values.

# Inspiration
# p1 <- ggplot(aes(x = age, y = friend_count_mean), 
#       data = subset(pf.fc_by_age, age < 71)) 

# Inspiration
# ggplot(aes(x = www_likes_received, y = likes_received), data = pf) + geom_point() +
#  xlim(0, quantile(pf$www_likes_received, 0.95)) +
#  ylim(0, quantile(pf$likes_received, 0.95))


# Create scatterplot - using GGPlot, change the transparancy
ggplot(diamonds, aes(carat, price)) +
  geom_jitter(alpha = 1/30) +
  xlim(0, quantile(diamonds$carat, 0.99)) +
  ylim(0, quantile(diamonds$price, 0.99))
```
```{r}
# What is the max carat size
max(diamonds[,1])
# What is the max price
max(diamonds[,7])
```
## Price vs. volume
```{r}
# Create a scatterplot of price vs. volume (x * y * z).
# This is a very rough approximation for a diamond's volume.

# Create a new variable for volume in the diamonds data frame.
# This will be useful in a later exercise.

# Don't make any adjustments to the plot just yet.

# This assignment is not graded and
# will be marked as correct when you submit.

diamonds$volume <- diamonds$x*diamonds$y*diamonds$z
summary(diamonds)
ggplot(diamonds, aes(volume, price)) + 
  geom_point()
```
## Correlations on Subsets
```{r}
diamonds0800 <- subset(diamonds, (volume > 0)&(volume < 800) )
with(diamonds0800, cor.test(volume, price))
```
## Adjustments - Price vs. volume
```{r}
# Subset the data to exclude diamonds with a volume
# greater than or equal to 800. Also, exclude diamonds
# with a volume of 0. Adjust the transparency of the
# points and add a linear model to the plot. (See the
# Instructor Notes or look up the documentation of
# geom_smooth() for more details about smoothers.)

# We encourage you to think about this next question and
# to post your thoughts in the discussion section.

# Do you think this would be a useful model to estimate
# the price of diamonds? Why or why not?

# This assignment is not graded and
# will be marked as correct when you submit.
ggplot(diamonds0800, aes(volume, price)) + 
  geom_jitter(alpha = 1/30) + 
  geom_smooth()
```


Do you think this would be a useful model to estimate the price of diamonds? Why or why not?

Does the linear model seem to be a good fit to the data? Share your thoughts.

It seems to be a good model as it fits quite well to the area

##Mean Price by Clarity

```{r}
# Use the function dplyr package
# to create a new data frame containing
# info on diamonds by clarity.

# Name the data frame diamondsByClarity

# The data frame should contain the following
# variables in this order.

#       (1) mean_price
#       (2) median_price
#       (3) min_price
#       (4) max_price
#       (5) n

# where n is the number of diamonds in each
# level of clarity.

# This assignment WILL BE automatically
# graded!
library(dplyr)

diamondsByClarity <- group_by(diamonds, clarity) %>%
  summarise( mean_price=mean(price),
             median_price=median(price),
             min_price=min(price),
             max_price=max(price),
             n=n())

head(diamondsByClarity)
```

## Bar Chart of Mean Prices
```{r}
# We’ve created summary data frames with the mean price
# by clarity and color. You can run the code in R to
# verify what data is in the variables diamonds_mp_by_clarity
# and diamonds_mp_by_color.

# Your task is to write additional code to create two bar plots
# on one output image using the grid.arrange() function from the package
# gridExtra.

# This assignment is not graded and
# will be marked as correct when you submit.

# See the Instructor Notes for more info on bar charts
# and for a hint on this task.

# DO NOT DELETE THE LINES OF CODE BELOW
# ===================================================================
data(diamonds)
library(dplyr)

diamonds_by_clarity <- group_by(diamonds, clarity)
diamonds_mp_by_clarity <- summarise(diamonds_by_clarity, mean_price = mean(price))

diamonds_by_color <- group_by(diamonds, color)
diamonds_mp_by_color <- summarise(diamonds_by_color, mean_price = mean(price))
```

```{r}
library(gridExtra)

p1 <- ggplot(diamonds_mp_by_clarity, aes(clarity, mean_price)) + 
  geom_bar(stat = 'identity')
p2 <- ggplot(diamonds_mp_by_color, aes(color, mean_price)) + 
  geom_bar(stat = 'identity')

grid.arrange(p1, p2, ncol = 1)
```
Mean price tends to decrease as clarity improves. The same can be said for color.
```{r}
diamonds_by_cut <- group_by(diamonds, cut)
diamonds_mp_by_cut <- summarise(diamonds_by_cut, mean_price = mean(price))
p3 <- ggplot(diamonds_mp_by_cut, aes(cut, mean_price)) + 
  geom_bar(stat = 'identity')

grid.arrange(p1, p2, p3, ncol = 1)
```


