Lesson 3
==============================================

## Reading in Data
```{r}
getwd()
list.files()

```

```{r}
setwd("C:/Users/Serg/Documents/Aprendiendo/Data Analysis/Module 5 - Exploratory Data Analysis (EDA)/Material del curso/lesson3")
list.files()
```

```{r}
pf <- read.csv("pseudo_facebook.tsv", sep = "\t")
names(pf)
```

## Histogram of Users' Birthdays

```{r}
install.packages("ggplot2")
library(ggplot2)

names(pf)
ggplot(data = pf, aes(x = dob_day)) +
  geom_histogram(binwidth = 1) +
  scale_x_continuous(breaks = 1:31)
```

## Moira's experiment

```{r}
ggplot(data = pf, aes(x = dob_day)) +
  geom_histogram(binwidth = 1) +
  scale_x_continuous(breaks = 1:31) +
  facet_wrap(~dob_month, ncol = 3 )
```

## Friend count

```{r}
qplot(data = pf, x = friend_count)
```

## Limiting the Axes
```{r}
qplot(data = pf, x = friend_count, xlim = c(0, 1000))

qplot(data = pf, x = friend_count) + 
  scale_x_continuous(limits = c(0, 1000))

```

## Adjusting the Bin Width
```{r}
qplot(data = pf,x = friend_count, binwidth = 25) +
  scale_x_continuous(limits = c(0, 1000), breaks = seq(0, 1000, 50)) +
  facet_wrap(~gender)
```

## Omitting NA Gender Observations
```{r}
qplot(data = subset(pf, !is.na(gender)),x = friend_count, binwidth = 25) +
  scale_x_continuous(limits = c(0, 1000), breaks = seq(0, 1000, 50)) +
  facet_wrap(~gender)
```
## Omitting ANY NA Observations in the Dataset
```{r}
qplot(data = na.omit(pf),x = friend_count, binwidth = 25) +
  scale_x_continuous(lim = c(0, 1000), breaks = seq(0, 1000, 50)) +
  facet_wrap(~gender)

# The na.omit wrapper removes any rows (or users) that have NA for ANY of te variables in the dataset
```

## Statistics 'by' gender
```{r}
table(pf$gender)

by(pf$friend_count, pf$gender, summary)
```

## Tenure
```{r}
qplot(x = tenure, data = pf, binwidth = 30,
      color = I('black'), fill = I('#099DD9'))

```

## Tenure solution
```{r}
qplot(x = tenure/365, data = pf, binwidth = .25,
      color = I('black'), fill = I('#F79420')) +
  scale_x_continuous(breaks = seq(1, 7, 1), lim = c(0, 7))

```

##] Labeling Plots
```{r}
qplot(x = tenure/365, data = pf, binwidth = .25,
      xlab = 'Number of years using Facebook',
      ylab = 'Number of users in sample',
      color = I('black'), fill = I('#F79420')) +
  scale_x_continuous(breaks = seq(1, 7, 1), lim = c(0, 7))
```

## User Ages
```{r}
qplot(x = age, data = pf, binwidth = 1,
      color = I('black'), fill = (I('#AAAAAA'))) + 
  scale_x_continuous(breaks = seq(0, 113, 5), lim = c(0, 113))
```

## Transforming Data
```{r}
qplot(x = friend_count, data = pf)

summary(pf$friend_count)

summary(log10(pf$friend_count + 1))

summary(sqrt(pf$friend_count))
```

## In order to visualize more than on histogram in the same plot
```{r}
install.packages('gridExtra')
library(gridExtra)
```

## Generate individual plots and arrange in grid
```{r}
p1 <- qplot(x = friend_count, data = pf)

p2 <- qplot(x = log10(friend_count + 1), data = pf)
  
p3 <- qplot(x = sqrt(friend_count), data = pf)


#arrange plots in grid

grid.arrange(p1, p2, p3, ncol = 3)
```

## Transforming Data Alternative Solution
```{r}
p1 <- ggplot(aes(x = friend_count), data = pf) + geom_histogram()
p2 <- p1 + scale_x_log10()
p3 <- p1 + scale_x_sqrt()
grid.arrange(p1, p2, p3, ncol = 1)

```

## Add a scaling layer
```{r}
logScale <- qplot(x = log10(friend_count + 1), data = pf)

countScale <- ggplot(aes(x = friend_count), data = pf) +
  geom_histogram() + 
  scale_x_log10()

grid.arrange(logScale, countScale, ncol = 2)

qplot(data = pf, x = friend_count) +
  scale_x_log10()
```

## Frequency Polygons (before we had histograms)
```{r}
qplot(x = friend_count, data = subset(pf, !is.na(gender)),
      binwidth = 10) +
  scale_x_continuous(lim = c(0, 1000), breaks = seq(0, 1000, 50)) +
  facet_wrap(~gender)

qplot(x = friend_count, y = ..count../sum(..count..), 
      data = subset(pf, !is.na(gender)),
      xlab = 'Friend Count',
      ylab = 'Proportion of Users with that friend count',
      binwidth = 10, geom = 'freqpoly', color = gender) +
  scale_x_continuous(lim = c(0, 1000), breaks = seq(0, 1000, 50)) 
```

## Frequency Polygons Solution
```{r}
qplot(x = www_likes, data = subset(pf, !is.na(gender)),
      geom = 'freqpoly', color = gender) + 
  scale_x_log10()
```

## Likes on the Web
```{r}
by(pf$www_likes, pf$gender, sum)
```

## Histograms
```{r}
qplot(x = friend_count, data = subset(pf, !is.na(gender)),
      binwidth = 25) +
  scale_x_continuous(lim = c(0, 1000), breaks = seq(0, 1000, 50)) +
  facet_wrap(~gender)
```

## Box Plots
```{r}
qplot(x = gender, y = friend_count, 
      data = subset(pf,!is.na(gender)),
      geom = 'boxplot', ylim = c(0, 500))

qplot(x = gender, y = friend_count, 
      data = subset(pf,!is.na(gender)),
      geom = 'boxplot') +
  scale_y_continuous(lim = c(0, 500))

qplot(x = gender, y = friend_count, 
      data = subset(pf,!is.na(gender)),
      geom = 'boxplot') +
  coord_cartesian(ylim = c(0, 500))
```

## Box Plots, Quartiles, and Friend Requests
```{r}
qplot(x = gender, y = friend_count,
      data = subset(pf, !is.na(gender)),
      geom = 'boxplot') + 
  coord_cartesian(ylim = c(0, 250))

by(pf$friend_count, pf$gender, summary)
```

## Box Plots, Quartiles, and Friend requests solution

Who made more friend requests - men or women?
```{r}
qplot(x = gender, y = friendships_initiated,
      data = subset(pf, !is.na(gender)), geom = 'boxplot') +
  coord_cartesian(ylim = c(0, 150))

by(pf$friendships_initiated, pf$gender, summary)
```

## Getting Logical
```{r}
summary(pf$mobile_likes)

summary(pf$mobile_likes > 0)

pf$mobile_check_in <- NA
pf$mobile_check_in <- ifelse(pf$mobile_likes > 0, 1, 0)
pf$mobile_check_in <- factor(pf$mobile_check_in)
summary(pf$mobile_check_in)

```

## Getting Logical solution
```{r}
summary(pf$mobile_check_in)
sum(pf$mobile_check_in == 1)/length(pf$mobile_check_in)
```

